### 2024-04-02 paper/paper.qmd
initial setup

---
title: "Decoding Sleep Disorders Through Self-Reported Patterns"
subtitle: "A Logistic Regression Approach to the NHANES 2017-March 2020 Sleep Data"
title-block-banner: true
abstract: "This study investigates the potential relationship between individuals' sleep patterns and the self-reported incidence of sleep troubles as confirmed by a medical professional. Using logistic regression analysis on data from the NHANES 2017-March 2020 Sleep Disorders dataset, we examine how various factors such as sleep duration on weekdays and weekends, frequency of snoring, and daytime sleepiness are associated with the likelihood of reporting sleep problems to a doctor. The findings aim to shed light on the predictive value of self-monitored sleep behaviors in identifying individuals who may require medical attention for sleep-related issues. This analysis contributes to the broader understanding of sleep health and its complex interactions with daily functioning."
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Sleep-Disorder-Analysis"
author: "Terry Tu"
date: today
date-format: long
format: pdf
toc: true
toc-depth: 1
number-sections: true
bibliography: references.bib
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(haven)
library(arrow)
```

# Introduction

The rhythm of our nightly rest is more than a personal habit; it's a public health pulse that captures the essence of well-being in our fast-paced society. The increasing prevalence of sleep disorders and their impact on daily life and overall health has become a point of societal concern, akin to the growing conversation around mental health and lifestyle diseases. The intimate link between quality of sleep and the vitality of an individual's life prompts a closer examination of sleep patterns within the populace. Drawing on the rich dataset from the National Health and Nutrition Examination Survey (NHANES) for the years 2017 to March 2020, our research delves into the self-reported instances of sleep disturbances and their association with various sleep behaviors.

This study is an explorative journey into the silent epidemic of sleep disorders that plague modern society, affecting productivity, mental health, and long-term wellbeing. The aim is to uncover the underlying patterns of sleep behavior that correlate with the reports of sleep troubles to medical professionals, thereby piecing together the nocturnal puzzle of restless societies. We explore the quantitative relationship between self-reported snoring frequency, feelings of daytime sleepiness, and the regularity of sleep hours during weekdays and weekends with the likelihood of reporting sleep issues to a doctor [@Prevention_2021].

Our analysis hinges on the application of logistic regression to the NHANES dataset, which presents a comprehensive view of American sleep habits. By interpreting the nuances of this rich dataset, we aspire to illuminate the factors that signal the need for medical attention in the domain of sleep health. The outcome of our investigation is poised to provide a scaffold for healthcare professionals and policymakers to base early intervention strategies, aiming to cultivate a well-rested population.

Following this introduction, the structure of the paper is laid out to facilitate a coherent flow of information and analysis. @sec-data (Data) provides a meticulous breakdown of the NHANES dataset, elucidating the data cleaning process and offering a descriptive overview of the key variables. @sec-model (Model) details the logistic regression model's design and the rationale behind the choice of predictors. @sec-result (Result) presents the findings, interpreted with precision and caution, alongside graphical representations for clarity. Concluding the paper, @sec-discussion (Discussion) reflects on the broader implications of the study, acknowledging limitations and proposing avenues for future research.

# Data {#sec-data}

```{r include=FALSE, message=FALSE}
# read the data file and store it in the 'data' variable
cleaned_data <- read_parquet(here("data", "analysis_data", "cleaned_data.parquet"))
```

## Raw Data

The dataset underpinning this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020. This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially. After meticulous data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional. The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].

The survey participants' ages range widely, reflecting the diversity of the American population. Variables are finely tuned to capture the multifaceted nature of sleep, encompassing aspects such as duration, disruptions, and subjective experiences of daytime sleepiness. The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines. For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-appendix (Appendix).

## Data Analysis Tools

Our statistical exploration was conducted within the R programming environment [@r], leveraging its comprehensive ecosystem for data analysis. We utilized the tidyverse collection of R packages [@rTidyverse] to streamline our data processing tasks. The ggplot2 package [@rGgplot2] was instrumental in crafting insightful visualizations that articulated the intricate relationships within our data. The dplyr package [@rDplyr] provided a syntax that facilitated the manipulation and transformation of our dataset, enabling us to prepare the data effectively for logistic regression analysis. Data importation was efficiently handled by the readr package [@rReadr], known for its quick and user-friendly approach to reading tabular data. Navigational simplicity within our project's directories was achieved with the here package [@R-here], which reliably managed file paths without the need for manual path setting. The reproducibility of our research was enhanced by the knitr package [@rKnitr], which seamlessly wove R code into our report, ensuring that our findings are transparent and replicable. For tabular data presentation, kableExtra[@R-kableExtra] offered a suite of customization options that enhanced the readability and aesthetic appeal of our tables. The logistic regression model was developed using core functions in R, which provide robust methods for estimating the effects of various predictors on a binary outcome.

## Variable Description {#sec-variable}

**Weekday Sleep Duration (SLD012)**: This variable measures the total number of hours respondents usually sleep on weekdays or workdays. It is derived from their reported usual bedtime and wake time, providing an insight into their sleep patterns during the typical workweek.

**Weekend Sleep Duration (SLD013)**: Similar to the weekday sleep duration, this variable represents the total number of hours respondents usually sleep on weekends or non-workdays. It helps in understanding the variation in sleep patterns during days off from work.

**Snoring Frequency (SLQ030)**: This variable records how often respondents snore while sleeping, with responses ranging from 'Never' to 'Frequently'. Snoring is a common symptom of sleep disorders such as obstructive sleep apnea, making this variable relevant to the study of sleep health.

**Daytime Sleepiness (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day. It is an indicator of sleep quality and quantity, as well as potential sleep disorders.

**Reported Sleep Trouble (SLQ050)**: The dependent variable in this study, it indicates whether respondents have ever told a doctor or other health professional that they have trouble sleeping. It is treated as a binary outcome variable, where '1' signifies reporting sleep trouble and '2' signifies no report of sleep trouble.

# Model {#sec-model}

# Result {#sec-result}

# Discussion {#sec-discussion}

# Appendix {#sec-appendix}

# References

### 2024-04-02 scripts/02_cleanup-data.R

@@ -1,0 +1,43 @@
+#### Preamble ####
+# Purpose: To clean the NHANES sleep disorder data by renaming columns for better understanding and saving the data as a parquet file
+# Author: Terry Tu
+# Date: 2 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'readr', 'dplyr', 'here', 'haven', and 'arrow' for parquet files
+# The raw dataset must be located in the 'data/raw_data' directory before running this script
+
+#### Workspace setup ####
+library(readr)
+library(dplyr)
+library(here)
+library(haven)
+library(arrow)
+
+#### Data Import ####
+data <- read_xpt(here("data", "raw_data", "P_SLQ.XPT"))
+
+#### Data Cleaning ####
+# Rename columns for clarity
+cleaned_data <- data %>%
+  rename(
+    RespondentID = SEQN,
+    WeekdaySleepTime = SLQ300,
+    WeekdayWakeTime = SLQ310,
+    WeekdaySleepDuration = SLD012,
+    WeekendSleepTime = SLQ320,
+    WeekendWakeTime = SLQ330,
+    WeekendSleepDuration = SLD013,
+    SnoringFrequency = SLQ030,
+    BreathingPauseFrequency = SLQ040,
+    ReportedSleepTrouble = SLQ050,
+    DaytimeSleepiness = SLQ120
+  ) %>%
+  # Remove NA values, if necessary (depending on the analysis you're doing)
+  filter(complete.cases(.))
+
+#### Data Export ####
+# Save the cleaned data as CSV
+write_csv(cleaned_data, here("data", "analysis_data", "cleaned_data.csv"))
+# Save the cleaned data as a parquet file
+write_parquet(cleaned_data, here("data", "analysis_data", "cleaned_data.parquet"))

### 2024-04-11 scripts/02_cleanup-data.R
Fixed cleanup script and added the test script

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: To clean the NHANES sleep disorder data by renaming columns for better understanding and saving the data as a parquet file
+# Purpose: To clean the NHANES sleep disorder data by renaming columns, removing missing/refused/don't know responses, removing unnecessary columns, and saving the data as a parquet file
 # Author: Terry Tu
 # Date: 2 April 2024
 # Contact: xiangyu.tu@mail.utoronto.ca
@@ -18,26 +18,38 @@ library(arrow)
 data <- read_xpt(here("data", "raw_data", "P_SLQ.XPT"))
 
 #### Data Cleaning ####
-# Rename columns for clarity
+# Rename columns for clarity, remove missing/refused/don't know responses, and select only the relevant columns
 cleaned_data <- data %>%
   rename(
     RespondentID = SEQN,
-    WeekdaySleepTime = SLQ300,
-    WeekdayWakeTime = SLQ310,
     WeekdaySleepDuration = SLD012,
-    WeekendSleepTime = SLQ320,
-    WeekendWakeTime = SLQ330,
     WeekendSleepDuration = SLD013,
     SnoringFrequency = SLQ030,
     BreathingPauseFrequency = SLQ040,
     ReportedSleepTrouble = SLQ050,
-    DaytimeSleepiness = SLQ120
+    OverlySleepFrequency = SLQ120
   ) %>%
-  # Remove NA values, if necessary (depending on the analysis you're doing)
-  filter(complete.cases(.))
+  filter(
+    !is.na(WeekdaySleepDuration),
+    !is.na(WeekendSleepDuration),
+    !is.na(SnoringFrequency) & SnoringFrequency != 7 & SnoringFrequency != 9,
+    !is.na(BreathingPauseFrequency) & BreathingPauseFrequency != 7 & BreathingPauseFrequency != 9,
+    !is.na(ReportedSleepTrouble) & ReportedSleepTrouble != 7 & ReportedSleepTrouble != 9,
+    !is.na(OverlySleepFrequency) & OverlySleepFrequency != 7 & OverlySleepFrequency != 9
+  ) %>%
+  select(
+    RespondentID,
+    WeekdaySleepDuration,
+    WeekendSleepDuration,
+    SnoringFrequency,
+    BreathingPauseFrequency,
+    ReportedSleepTrouble,
+    OverlySleepFrequency
+  )
 
 #### Data Export ####
 # Save the cleaned data as CSV
[…]

### 2024-04-11 scripts/03_test-data-validity.R

@@ -1,0 +1,53 @@
+#### Preamble ####
+# Purpose: To perform a series of data integrity tests on the cleaned NHANES sleep disorder data
+# Author: Terry Tu
+# Date: 2 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'dplyr', 'here', 'readr', 'testthat'
+# The cleaned dataset must be located in the 'data/analysis_data' directory before running this script
+
+#### Workspace setup ####
+library(readr)
+library(testthat)
+library(dplyr)
+library(here)
+
+# Load the cleaned data
+cleaned_data <- read_csv(here("data", "analysis_data", "cleaned_data.csv"))
+
+# Test 1: Check if every column has a valid integer ID (SEQN)
+test_that("Every column has a valid integer ID", {
+  expect_true(all(!is.na(cleaned_data$RespondentID) & cleaned_data$RespondentID == as.integer(cleaned_data$RespondentID)))
+})
+
+# Test 2: Check for valid values in WeekdaySleepDuration (SLD012)
+test_that("Valid values in WeekdaySleepDuration", {
+  expect_true(all(!is.na(cleaned_data$WeekdaySleepDuration) & cleaned_data$WeekdaySleepDuration >= 2 & cleaned_data$WeekdaySleepDuration <= 14))
+})
+
+# Test 3: Check for valid values in WeekendSleepDuration (SLD013)
+test_that("Valid values in WeekendSleepDuration", {
+  expect_true(all(!is.na(cleaned_data$WeekendSleepDuration) & cleaned_data$WeekendSleepDuration >= 2 & cleaned_data$WeekendSleepDuration <= 14))
+})
+
+# Test 4: Check for valid values in SnoringFrequency (SLQ030)
+test_that("Valid values in SnoringFrequency", {
+  expect_true(all(!is.na(cleaned_data$SnoringFrequency) & cleaned_data$SnoringFrequency %in% 0:3))
+})
+
+# Test 5: Check for valid values in BreathingPauseFrequency (SLQ040)
+test_that("Valid values in BreathingPauseFrequency", {
+  expect_true(all(!is.na(cleaned_data$BreathingPauseFrequency) & cleaned_data$BreathingPauseFrequency %in% 0:3))
+})
+
+# Test 6: Check for valid values in ReportedSleepTrouble (SLQ050)
[…]

### 2024-04-12 paper/paper.qmd
added logistic model

@@
 
-**Weekday Sleep Duration (SLD012)**: This variable measures the total number of hours respondents usually sleep on weekdays or workdays.
-It is derived from their reported usual bedtime and wake time, providing an insight into their sleep patterns during the typical workweek.
+**Weekday Sleep Duration (SLD012)**: This variable measures the total number of hours respondents usually sleep on weekdays or workdays, with values ranging from 2 to 14 hours.
+It provides insight into their sleep patterns during the typical workweek.
 
-**Weekend Sleep Duration (SLD013)**: Similar to the weekday sleep duration, this variable represents the total number of hours respondents usually sleep on weekends or non-workdays.
+**Weekend Sleep Duration (SLD013)**: Similar to the weekday sleep duration, this variable represents the total number of hours respondents usually sleep on weekends or non-workdays, also ranging from 2 to 14 hours.
 It helps in understanding the variation in sleep patterns during days off from work.
 
-**Snoring Frequency (SLQ030)**: This variable records how often respondents snore while sleeping, with responses ranging from 'Never' to 'Frequently'.
+**Snoring Frequency (SLQ030)**: This variable records how often respondents snore while sleeping, with responses ranging from 0 (Never) to 3 (Frequently).
 Snoring is a common symptom of sleep disorders such as obstructive sleep apnea, making this variable relevant to the study of sleep health.
 
-**Daytime Sleepiness (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day.
+**Daytime Sleepiness (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day, with values ranging from 0 (Never) to 4 (Almost always).
 It is an indicator of sleep quality and quantity, as well as potential sleep disorders.
@@
 **Reported Sleep Trouble (SLQ050)**: The dependent variable in this study, it indicates whether respondents have ever told a doctor or other health professional that they have trouble sleeping.
-It is treated as a binary outcome variable, where '1' signifies reporting sleep trouble and '2' signifies no report of sleep trouble.
+It is treated as a binary outcome variable, with values of 0 (No report of sleep trouble) and 1 (Reporting sleep trouble).
 
 # Model {#sec-model}
+
+The aim of our model is to explore the relationship between various sleep-related factors and the self-reported incidence of sleep troubles.
+We employ a Bayesian logistic regression model to analyze the data from the National Health and Nutrition Examination Survey (NHANES).
+Further details and diagnostics of this model are provided in @sec-posterior.
+
+## Model set-up
+
+Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise.
+The predictors include weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$).
+The logistic regression model is formulated as follows:
+
+\begin{align}
+y_i | p_i &\sim \mbox{Bernoulli}(p_i) \\
+\mbox{logit}(p_i) &= \beta_0 + \beta_1 \times x_{i1} + \beta_2 \times x_{i2} + \beta_3 \times x_{i3} + \beta_4 \times x_{i4} + \beta_5 \times x_{i5} \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1, \beta_2, \beta_3, \beta_4, \beta_5 &\sim \mbox{Normal}(0, 2.5)
+\end{align}
+
+The model is implemented in R using the rstanarm package, enabling the incorporation of prior beliefs about the parameters and estimation of their posterior distributions.
+The stan_glm function from rstanarm is used to perform the Bayesian logistic regression, integrating both the data and prior knowledge for inference.
+
+Priors for the intercept and coefficients are chosen to reflect a neutral stance before observing the data, centered around zero with a moderate spread to allow flexibility in the estimates.
+These priors convey an expectation of no substantial effect from each predictor, with the data informing deviations from this baseline assumption.
+
+The logistic regression framework models the log odds of reporting sleep troubles as a linear combination of the predictors, with the coefficients indicating the direction and magnitude of each predictor's effect.
+The stan_glm function combines these priors with the data to estimate the posterior distributions of the model parameters, providing a comprehensive view of the relationships between sleep-related factors and reported sleep troubles.
+
+## Model Justification
+In our analysis, we hypothesize that certain sleep-related factors will have significant associations with the likelihood of reporting sleep troubles. Specifically:
+
+**Weekday Sleep Duration ($\beta_1$)**: We hypothesize that this coefficient may be positive, indicating that longer sleep durations on weekdays could be associated with a higher likelihood of reporting sleep troubles.
+This could reflect the possibility that individuals with sleep disturbances may attempt to compensate by sleeping longer.
+
+**Weekend Sleep Duration ($\beta_2$)**: Conversely, we hypothesize that this coefficient may be negative, suggesting that longer sleep durations on weekends might be associated with a lower likelihood of reporting sleep troubles.
+This could indicate that catching up on sleep during weekends may alleviate some sleep-related issues.
+
+**Snoring Frequency ($\beta_3$)**: We expect this coefficient to be positive, as frequent snoring is often a symptom of sleep disorders such as obstructive sleep apnea, which could lead to reporting sleep troubles.
+
+**Breathing Pause Frequency ($\beta_4$)**: Similarly, we hypothesize that this coefficient will be positive, as pauses in breathing during sleep are a hallmark of sleep apnea, a condition commonly associated with sleep disturbances.
+
+**Daytime Sleepiness ($\beta_5$)**: We anticipate that this coefficient will also be positive, as excessive sleepiness during the day is a common consequence of poor sleep quality or insufficient sleep, which may prompt individuals to report sleep troubles.
+
+The Bayesian framework allows us to incorporate these hypotheses as priors in our model, providing a structured way to integrate our substantive expectations with the data analysis.
+By doing so, we can obtain a more informed and nuanced interpretation of the relationships between sleep-related factors and the likelihood of reporting sleep troubles.
 

### 2024-04-12 scripts/02_cleanup-data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: To clean the NHANES sleep disorder data by renaming columns, removing missing/refused/don't know responses, removing unnecessary columns, and saving the data as a parquet file
+# Purpose: To clean the NHANES sleep disorder data by renaming columns, removing missing/refused/don't know responses, removing unnecessary columns, mutating values, and saving the data as a parquet file
 # Author: Terry Tu
 # Date: 2 April 2024
 # Contact: xiangyu.tu@mail.utoronto.ca
@@ -18,7 +18,7 @@ library(arrow)
 data <- read_xpt(here("data", "raw_data", "P_SLQ.XPT"))
 
 #### Data Cleaning ####
-# Rename columns for clarity, remove missing/refused/don't know responses, and select only the relevant columns
+# Rename columns for clarity, remove missing/refused/don't know responses, select only the relevant columns, and mutate ReportedSleepTrouble from 2 to 0
 cleaned_data <- data %>%
   rename(
     RespondentID = SEQN,
@@ -37,6 +37,7 @@ cleaned_data <- data %>%
     !is.na(ReportedSleepTrouble) & ReportedSleepTrouble != 7 & ReportedSleepTrouble != 9,
     !is.na(OverlySleepFrequency) & OverlySleepFrequency != 7 & OverlySleepFrequency != 9
   ) %>%
+  mutate(ReportedSleepTrouble = ifelse(ReportedSleepTrouble == 2, 0, ReportedSleepTrouble)) %>%
   select(
     RespondentID,
     WeekdaySleepDuration,
@@ -52,4 +53,3 @@ cleaned_data <- data %>%
 write_csv(cleaned_data, here("data", "analysis_data", "cleaned_data.csv"))
 # Save the cleaned data as a parquet file
 write_parquet(cleaned_data, here("data", "analysis_data", "cleaned_data.parquet"))
-

### 2024-04-12 scripts/03_test-data-validity.R

@@ -43,7 +43,7 @@ test_that("Valid values in BreathingPauseFrequency", {
 
 # Test 6: Check for valid values in ReportedSleepTrouble (SLQ050)
 test_that("Valid values in ReportedSleepTrouble", {
-  expect_true(all(!is.na(cleaned_data$ReportedSleepTrouble) & cleaned_data$ReportedSleepTrouble %in% c(1, 2)))
+  expect_true(all(!is.na(cleaned_data$ReportedSleepTrouble) & cleaned_data$ReportedSleepTrouble %in% c(0, 1)))
 })
 
 # Test 7: Check for valid values in OverlySleepFrequency (SLQ120)

### 2024-04-12 scripts/04_model.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: To construct a Bayesian logistic regression model assessing the impact of sleep-related variables on reported sleep trouble using rstanarm
+# Author: Terry Tu
+# Date: 2 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The 'rstanarm' package for Bayesian modeling, 'readr' for data manipulation, and 'here' for path management
+# The 'cleaned_data' dataset must be pre-processed before using it in the model.
+
+#### Workspace setup ####
+library(readr)
+library(rstanarm)
+library(here)
+
+#### Read data ####
+cleaned_data <- read_csv(here("data", "analysis_data", "cleaned_data.csv"))
+
+# Ensure reproducibility
+set.seed(2024)
+
+#### Model data ####
+bayesian_logistic_model <- stan_glm(
+  ReportedSleepTrouble ~ WeekdaySleepDuration + WeekendSleepDuration + SnoringFrequency + BreathingPauseFrequency + OverlySleepFrequency,
+  data = cleaned_data,
+  family = binomial(), # Indicating a logistic regression model
+  prior = normal(0, 2.5), # Set the prior for the regression coefficients
+  prior_intercept = normal(0, 2.5), # Set the prior for the intercept
+  seed = 2024 # Set the seed for reproducibility
+)
+
+# Summary of the model to check for immediate red flags
+summary(bayesian_logistic_model)
+
+#### Save model ####
+# Save the Bayesian logistic regression model as an .rds file to the 'models' folder
+saveRDS(
+  bayesian_logistic_model,
+  file = here("models", "bayesian_logistic_model.rds")
+)

### 2024-04-13 paper/paper.qmd
added measurement

@@
 
+## Measurement:
+
+In this study, we utilized data from the National Health and Nutrition Examination Survey (NHANES), specifically focusing on the sleep disorders component which includes data collected between 2017 and March 2020.
+The NHANES program, a longstanding project conducted by the National Center for Health Statistics (NCHS), plays a critical role in assessing the health and nutritional status of adults and children in the United States.
+This dataset is pivotal in understanding public health and informs policy decisions through scientifically reliable data [@Prevention_2021].
+
+The sleep disorders dataset within NHANES is enriched by questions adapted from the Munich ChronoType Questionnaire [@Roenneberg2003], targeting various aspects of sleep behavior and disorders.
+The inclusion of these questions is instrumental in exploring the complex dynamics of sleep patterns among the U.S.
+population.
+Due to disruptions caused by the COVID-19 pandemic, the 2019-2020 data collection cycle was prematurely halted in March 2020, leading to its combination with the 2017-2018 cycle to ensure national representativeness and analytical robustness.
+
+This combined dataset referred to as the NHANES 2017-March 2020 pre-pandemic data, offers valuable insights into the sleep habits of Americans before the pandemic.
+It is instrumental for researchers and public health officials aiming to understand baseline sleep behaviors and potential disturbances across a broad demographic spectrum.
+
+To prepare the dataset for analysis, extensive data cleaning and processing were conducted.
+This included removing entries with missing, refused, or 'don't know' responses for critical variables such as sleep duration and frequency of snoring.
+Additionally, to address issues with data reliability and consistency, about 3% of the data underwent verification through audio recordings of the interviews.
+Moreover, for variables capturing sleep duration on weekdays (SLD012) and weekends (SLD013), reported times were meticulously reviewed, with outliers adjusted and rounded to the nearest half-hour, enhancing the data's accuracy and usability.
+
+Detailed descriptions of the variables used in this study, along with the specific adjustments made to the dataset, are available in  @sec-variable.
+This section is designed to provide a comprehensive understanding of the origins, processing, and analytical framework applied to each variable relevant to this study.
+
 # Model {#sec-model}
@@
 
-# Appendix {#sec-appendix}
+# Appendix {.unnumbered #sec-appendix}
+
+## Additional Data Details
+
+## Model Details {#sec-model-details}
+
+### Posterior Predictive Check {#sec-posterior}
 

### 2024-04-13 paper/paper.qmd
added sample data

@@
 The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
-For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-appendix (Appendix).
+For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-add-data.
 
@@
 
-**Daytime Sleepiness (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day, with values ranging from 0 (Never) to 4 (Almost always).
+**Overly Sleep Frequency (SLQ120)**: This variable assesses how often respondents feel excessively or overly sleepy during the day, with values ranging from 0 (Never) to 4 (Almost always).
 It is an indicator of sleep quality and quantity, as well as potential sleep disorders.
@@
 It is treated as a binary outcome variable, with values of 0 (No report of sleep trouble) and 1 (Reporting sleep trouble).
+
+## Sample of Cleaned Sleep Disorder Data
+
+```{r}
+#| label: tbl-sleep-data-sample
+#| tbl-cap: Sample of Sleep Disorder Data
+#| echo: false
+# Code created with the help from ChatGPT
+sample_data <- head(cleaned_data) %>%
+  select(RespondentID, WeekdaySleepDuration, WeekendSleepDuration, SnoringFrequency, BreathingPauseFrequency, OverlySleepFrequency, ReportedSleepTrouble)
+
+# Create a nicely formatted table with custom column names
+kable(sample_data, col.names = c("Respondent ID", "Weekday Sleep Duration (hrs)", "Weekend Sleep Duration (hrs)", "Snoring Frequency", "Breathing Pause Frequency", "Overly Sleep Frequency", "Reported Sleep Trouble"))
+
+```
+
+@tbl-sleep-data-sample represents a subset of the broader NHANES sleep disorder dataset.
+Each row in the table corresponds to an individual participant, uniquely identified by their Respondent ID.
+The "Weekday Sleep Duration (hrs)" and "Weekend Sleep Duration (hrs)" columns quantify the number of hours slept during the weekdays and weekends, respectively, providing a snapshot of the individual's sleep patterns.
+"Snoring Frequency" and "Breathing Pause Frequency" are categorical measures that reflect how often the respondents experience snoring and breathing pauses during sleep, common indicators of sleep disturbances such as sleep apnea.
+The "Overly Sleep Frequency" column indicates the frequency at which respondents report feeling overly sleepy during the day, a sign that can be indicative of inadequate sleep quality or quantity.
+Lastly, the "Reported Sleep Trouble" column is a binary measure showing whether the respondent has reported having sleep troubles to a health professional, with 0 signifying no reported trouble and 1 indicating reported trouble.
 
@@
 This section is designed to provide a comprehensive understanding of the origins, processing, and analytical framework applied to each variable relevant to this study.
+
+## Data Exploration: {#sec-exploration}
 
@@
 
-# Appendix {.unnumbered #sec-appendix}
+# Appendix {#sec-appendix}
 
-## Additional Data Details
+## Additional Data Details {#sec-add-data}
 
-## Model Details {#sec-model-details}
+## Model Details
 

### 2024-04-14 paper/paper.qmd
added new data plot

@@
 library(arrow)
+library(gridExtra)
 ```
@@
 
-```{r include=FALSE, message=FALSE}
+```{r include=FALSE, message=FALSE, echo=FALSE}
 # read the data file and store it in the 'data' variable
@@
 
+In this section, we explore the distributions of key variables related to sleep patterns and disorders in the NHANES dataset.
+Histograms provide visual insights into the frequency of reported sleep duration, snoring, breathing pauses, daytime sleepiness, and reported sleep troubles.
+
+```{r fig-histograms, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Distributions of sleep-related variables"}
+# Generate individual histograms for each variable
+p1 <- ggplot(cleaned_data, aes(x = WeekdaySleepDuration)) + 
+  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") + 
+  labs(x = "Weekday Sleep Duration (hrs)", y = "Count") +
+  theme_minimal()
+
+p2 <- ggplot(cleaned_data, aes(x = WeekendSleepDuration)) + 
+  geom_histogram(binwidth = 1, fill = "salmon", color = "black") + 
+  labs(x = "Weekend Sleep Duration (hrs)", y = "Count") +
+  theme_minimal()
+
+p3 <- ggplot(cleaned_data, aes(x = SnoringFrequency)) + 
+  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") + 
+  labs(x = "Snoring Frequency", y = "Count") +
+  theme_minimal()
+
+p4 <- ggplot(cleaned_data, aes(x = BreathingPauseFrequency)) + 
+  geom_histogram(binwidth = 1, fill = "plum", color = "black") + 
+  labs(x = "Breathing Pause Frequency", y = "Count") +
+  theme_minimal()
+
+p5 <- ggplot(cleaned_data, aes(x = OverlySleepFrequency)) + 
+  geom_histogram(binwidth = 1, fill = "gold", color = "black") + 
+  labs(x = "Overly Sleep Frequency", y = "Count") +
+  theme_minimal()
+
+p6 <- ggplot(cleaned_data, aes(x = as.factor(ReportedSleepTrouble))) + 
+  geom_bar(fill = "lightblue", color = "black") + 
+  labs(x = "Reported Sleep Trouble", y = "Count") +
+  scale_x_discrete(labels = c("0" = "No Trouble", "1" = "Trouble")) +
+  theme_minimal()
+
+# Arrange the plots into a 2x3 grid
+grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3)
+```
+
+**Weekday Sleep Duration (hrs)**: The histogram for weekday sleep duration exhibits a unimodal distribution, peaking around 7 to 8 hours, which aligns with the generally recommended sleep duration for adults.
+The symmetrical shape suggests a commonality in sleep patterns among individuals during the workweek.
+
+**Weekend Sleep Duration (hrs)**: Similarly unimodal, the weekend sleep duration distribution also peaks around the same range as weekday sleep but shows a noticeable tendency for slightly longer durations.
+This may suggest that individuals take the opportunity to sleep more on weekends, possibly compensating for the workweek.
+
+**Snoring Frequency**: The snoring frequency histogram reveals a broad distribution, with the majority of respondents indicating they never or rarely snore.
+A significant count of individuals also reports frequent snoring, suggesting that the experience of snoring is quite varied among the respondents.
+
+**Breathing Pause Frequency**: The distribution for breathing pause frequency is steeply skewed towards 'Never', suggesting that breathing pauses during sleep are infrequently experienced or reported by the majority of respondents.
+The scant instances of frequent breathing pauses may indicate a lower occurrence or a lack of self-awareness of these events.
+
+**Overly Sleep Frequency**: The distribution of daytime sleepiness frequency is right-skewed, with 'Sometimes (2-4 times a month)' emerging as the modal category.
+However, a notable proportion of respondents experience sleepiness during the day at least once a month, suggesting that daytime sleepiness is a common issue.
+
+**Reported Sleep Trouble**: The binary distribution for reported sleep trouble shows a larger proportion of individuals reporting no sleep trouble compared to those reporting trouble, approximately threefold.
+This ratio highlights that while sleep trouble is present, it is not reported by the majority of participants in this sample.
+
 # Model {#sec-model}

### 2024-04-14 paper/paper.qmd
added result

@@
 library(gridExtra)
+library(modelsummary)
 ```
@@
 \mbox{logit}(p_i) &= \beta_0 + \beta_1 \times x_{i1} + \beta_2 \times x_{i2} + \beta_3 \times x_{i3} + \beta_4 \times x_{i4} + \beta_5 \times x_{i5} \\
-\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_1, \beta_2, \beta_3, \beta_4, \beta_5 &\sim \mbox{Normal}(0, 2.5)
+\beta_0 &\sim \mbox{Normal}(0, 1) \\ % Assuming the log-odds of the base rate of reported sleep trouble is centered around 0
+\beta_1, \beta_2, \beta_3, \beta_4, \beta_5 &\sim \mbox{Normal}(0, 1) \\ % Weekday sleep duration effect and Weekend sleep duration effect, Snoring frequency,  Breathing pause frequency, and Overly sleep frequency
 \end{align}
 
-The model is implemented in R using the rstanarm package, enabling the incorporation of prior beliefs about the parameters and estimation of their posterior distributions.
-The stan_glm function from rstanarm is used to perform the Bayesian logistic regression, integrating both the data and prior knowledge for inference.
+The logistic regression model is defined using a Bayesian framework, implemented in R with the `rstanarm` package.
+This approach allows us to incorporate prior knowledge about the parameters and to estimate their posterior distributions based on the observed data.
 
-Priors for the intercept and coefficients are chosen to reflect a neutral stance before observing the data, centered around zero with a moderate spread to allow flexibility in the estimates.
-These priors convey an expectation of no substantial effect from each predictor, with the data informing deviations from this baseline assumption.
+In our model, the probability of an individual reporting sleep troubles, denoted by $p_i$, follows a Bernoulli distribution.
+The log odds of reporting sleep troubles are modeled as a linear combination of the predictors, with $\beta_0$ representing the intercept, and $\beta_1$ through $\beta_5$ representing the slopes for the respective predictors.
 
-The logistic regression framework models the log odds of reporting sleep troubles as a linear combination of the predictors, with the coefficients indicating the direction and magnitude of each predictor's effect.
-The stan_glm function combines these priors with the data to estimate the posterior distributions of the model parameters, providing a comprehensive view of the relationships between sleep-related factors and reported sleep troubles.
+The priors for the intercept ($\beta_0$) and the slopes ($\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$) are deliberately chosen to reflect initial neutrality or skepticism regarding the impact of each predictor on the likelihood of reporting sleep troubles.
+Specifically, $\beta_0$ is assigned a Normal(0, 1) prior, indicating our initial uncertainty about the base rate of sleep trouble in the population.
+
+For $\beta_1$ and $\beta_2$, which correspond to the effects of weekday and weekend sleep duration, we posit Normal(0, 0.5) priors.
+These priors are centered around zero, reflecting the number of sleep hour can be either positively or negatively contribute to sleep troubles, with a modest expectation of effect size.
+
+The coefficients $\beta_3$, $\beta_4$, $\beta_5$ relating to snoring frequency, breathing pause frequency, and overly sleep frequency are each given a Normal(0, 1) prior.
+These priors imply that, prior to analyzing the data, we do not have strong expectations about the direction or scale of these effects.
 
 ## Model Justification
-In our analysis, we hypothesize that certain sleep-related factors will have significant associations with the likelihood of reporting sleep troubles. Specifically:
 
-**Weekday Sleep Duration ($\beta_1$)**: We hypothesize that this coefficient may be positive, indicating that longer sleep durations on weekdays could be associated with a higher likelihood of reporting sleep troubles.
-This could reflect the possibility that individuals with sleep disturbances may attempt to compensate by sleeping longer.
+Our Bayesian logistic regression model is designed to investigate the associations between various sleep-related factors and the likelihood of reporting sleep troubles.
+The priors for the coefficients are set with specific hypotheses in mind, informed by existing literature and plausible biological mechanisms:
 
-**Weekend Sleep Duration ($\beta_2$)**: Conversely, we hypothesize that this coefficient may be negative, suggesting that longer sleep durations on weekends might be associated with a lower likelihood of reporting sleep troubles.
-This could indicate that catching up on sleep during weekends may alleviate some sleep-related issues.
+**Weekday Sleep Duration ($\beta_1$)**: We initially hypothesize that there may be a complex relationship between sleep duration on weekdays and sleep troubles.
+Long sleep durations could indicate an attempt to compensate for poor quality sleep, potentially related to sleep disturbances.
+As such, the prior for this coefficient is centered around zero, allowing the data to indicate the direction of association.
 
-**Snoring Frequency ($\beta_3$)**: We expect this coefficient to be positive, as frequent snoring is often a symptom of sleep disorders such as obstructive sleep apnea, which could lead to reporting sleep troubles.
+**Weekend Sleep Duration ($\beta_2$)**: For sleep duration on weekends, we consider the possibility of a recuperative effect, where catching up on sleep may reduce the likelihood of reporting troubles.
+This hypothesis is tentative, hence a prior centered around zero with a conservative standard deviation, permitting data-driven insights into this relationship.
 
-**Breathing Pause Frequency ($\beta_4$)**: Similarly, we hypothesize that this coefficient will be positive, as pauses in breathing during sleep are a hallmark of sleep apnea, a condition commonly associated with sleep disturbances.
+**Snoring Frequency ($\beta_3$)**: Frequent snoring is a recognized symptom of sleep disorders like obstructive sleep apnea, which can lead to disrupted sleep.
+We posit a positive association between snoring frequency and reported sleep troubles, reflected in the priors for this coefficient.
 
-**Daytime Sleepiness ($\beta_5$)**: We anticipate that this coefficient will also be positive, as excessive sleepiness during the day is a common consequence of poor sleep quality or insufficient sleep, which may prompt individuals to report sleep troubles.
+**Breathing Pause Frequency ($\beta_4$)**: Breathing pauses during sleep are indicative of sleep apnea, which is directly related to sleep troubles.
+We hypothesize a positive association for this coefficient, which is incorporated into the prior setting.
 
-The Bayesian framework allows us to incorporate these hypotheses as priors in our model, providing a structured way to integrate our substantive expectations with the data analysis.
-By doing so, we can obtain a more informed and nuanced interpretation of the relationships between sleep-related factors and the likelihood of reporting sleep troubles.
+**Daytime Sleepiness ($\beta_5$)**: Excessive daytime sleepiness is often the result of inadequate or disrupted nighttime sleep.
+We hypothesize a positive relationship between daytime sleepiness and the reporting of sleep troubles, expecting that those experiencing higher levels of sleepiness will be more likely to report troubles.
+
+These hypotheses are structured into our Bayesian model as informed priors, which, while reflecting our initial expectations, are sufficiently flexible to be shaped by the empirical data.
+This integration of prior beliefs and observed information will yield posterior distributions that convey a robust interpretation of how sleep behaviors relate to the reporting of sleep troubles.
 
 # Result {#sec-result}
+
+```{r echo=FALSE}
+# Retrieve the Bayesian logistic model from the saved .rds file
+bayesian_logistic_model <- readRDS(file = here("models", "bayesian_logistic_model.rds"))
+```
+
+## Model Coefficients Interpretation
+
+```{r}
+#| label: tbl-logistic-coefficients
+#| tbl-cap: "Summary Statistics for the Coefficients of the Logistic Model"
+#| echo: false
+# Round the coefficients directly when creating the data frame
+coefficients_df <- data.frame(
+  Term = names(coef(bayesian_logistic_model)),
+  Estimate = round(coef(bayesian_logistic_model), 3)
+)
+
+# Create the kable without row names
+coefficient_table <- kable(coefficients_df, 
+                           col.names = c("Term", "Estimate"), 
+                           align = 'c',
+                           row.names = FALSE)  # Ensure row names are not printed
+
+# Print the summary table
+coefficient_table
+```
+@tbl-logistic-coefficients presents the key coefficients for understanding the influence of various sleep-related variables on the likelihood of reporting sleep troubles:
+
+**Intercept ($\beta_0$)**:
+The intercept of the model, $\beta_0$, is estimated at -1.396.
+This value represents the log odds of reporting sleep troubles when all the predictors are held at zero.
+It suggests that, in the absence of any sleep issues or without considering any sleep behaviors, the baseline log odds of reporting sleep troubles to a doctor are negative, indicating a lower probability of reporting sleep troubles.
+
+**Weekday Sleep Duration ($\beta_1$)**:
+For the weekday sleep duration, the coefficient $\beta_1$ is 0.051, suggesting a slight increase in the log odds of reporting sleep troubles for each additional hour of sleep during the weekdays.
+This could be indicative of a situation where those who sleep more on weekdays might be doing so due to sleep issues that have led them to consult with a doctor.
+
+**Weekend Sleep Duration ($\beta_2$)**:
+Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131.
+This implies that for each additional hour of sleep during the weekends, there is a decrease in the log odds of reporting sleep troubles.
+It could be interpreted as those who manage to sleep more on weekends are less likely to report sleep problems, possibly because catching up on sleep may alleviate some of their weekday sleep deficits.
+
+**Snoring Frequency ($\beta_3$)**:
+The coefficient for snoring frequency, $\beta_3$, has a value of 0.032.
+While this is a positive value, suggesting that an increase in snoring frequency is associated with an increase in the likelihood of reporting sleep troubles, the effect size is relatively small.
+
+**Breathing Pause Frequency ($\beta_4$)**:
+The coefficient for breathing pause frequency, $\beta_4$, is 0.412, indicating a more substantial positive association with reporting sleep troubles.
+This aligns with clinical understanding, as breathing pauses are often associated with sleep disorders like sleep apnea, which can be a significant concern prompting medical consultation.
+
+**Overly Sleep Frequency ($\beta_5$)**:
+Lastly, the coefficient for overly sleep frequency, $\beta_5$, is 0.425, suggesting a strong relationship with the reporting of sleep troubles.
+This result is intuitive, as feeling excessively sleepy can be a direct symptom of poor sleep quality or a sleep disorder, leading to discussions with a healthcare provider.
+
+**Interpretation**:
+The model indicates that various factors are related to the reporting of sleep troubles.
+While longer weekday sleep duration slightly increases the log odds of reporting sleep troubles, longer weekend sleep can reduce it.
+Frequent snoring has a smaller effect compared to breathing pauses and daytime sleepiness, which both have significant positive associations with reporting sleep troubles.
+These findings underline the complexity of sleep behavior and its impact on sleep quality and health consultations.
+
+## Model Equation
+
+\begin{align}
+\text{logit}(p_i) &= -1.396 + 0.051 \times x_{i1} - 0.131 \times x_{i2} + 0.032 \times x_{i3} + 0.412 \times x_{i4} + 0.425 \times x_{i5} + \varepsilon_i
+\end{align}
 

### 2024-04-14 scripts/04_model.R

@@ -23,9 +23,8 @@ bayesian_logistic_model <- stan_glm(
   ReportedSleepTrouble ~ WeekdaySleepDuration + WeekendSleepDuration + SnoringFrequency + BreathingPauseFrequency + OverlySleepFrequency,
   data = cleaned_data,
   family = binomial(), # Indicating a logistic regression model
-  prior = normal(0, 2.5), # Set the prior for the regression coefficients
-  prior_intercept = normal(0, 2.5), # Set the prior for the intercept
-  seed = 2024 # Set the seed for reproducibility
+  prior = normal(0, 1), # Set the prior for the regression coefficients
+  prior_intercept = normal(0, 1), # Set the prior for the intercept
 )
 
 # Summary of the model to check for immediate red flags

### 2024-04-16 paper/paper.qmd
added results

@@
 library(modelsummary)
+library(bayesplot)
 ```
@@
 The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
-For further information on the data cleaning specifics and validation checks, please see the supplementary material in @sec-add-data.
+For further information on the data cleaning specifics and validation checks, please see the supplementary material in Appendix.
 
@@
 We employ a Bayesian logistic regression model to analyze the data from the National Health and Nutrition Examination Survey (NHANES).
-Further details and diagnostics of this model are provided in @sec-posterior.
+Further details and diagnostics of this model are provided in Appendix.
 
@@
 
+## Predict Sleep Disorder Through Self Reported Patterns
+
+This section provides a detailed comparison of predicted sleep trouble probabilities for two hypothetical individuals, Jack and Bob, based on their sleep behavior statistics.
+The table below illustrates the significant differences in their sleep patterns and how these patterns potentially influence the logistic regression model's predictions regarding their sleep troubles.
+
+```{r}
+#| label: tbl-sample-prediction
+#| tbl-cap: "Comparison of predictive factors for Jack and Bob"
+#| echo: false
+# Define the coefficients from the model
+intercept <- -1.396
+coef_weekday_sleep <- 0.051
+coef_weekend_sleep <- -0.131
+coef_snoring_frequency <- 0.032
+coef_breathing_pause <- 0.412
+coef_overly_sleep_frequency <- 0.425
+
+# Sample data for Jack and Bob
+data <- data.frame(
+  Name = c("Jack", "Bob"),
+  WeekdaySleep = c(5, 8),  # hours
+  WeekendSleep = c(8, 10),  # hours
+  SnoringFrequency = c(3, 0),  # 0: Never, 3: Frequently
+  BreathingPause = c(3, 0),  # 0: Never, 3: Frequently 
+  OverlySleepy = c(4, 1)  # 0: Never, 4: Always, 1: Rarely
+)
+
+# Calculate the logit(p) for each person
+data$logit_p <- intercept + 
+  coef_weekday_sleep * data$WeekdaySleep + 
+  coef_weekend_sleep * data$WeekendSleep + 
+  coef_snoring_frequency * data$SnoringFrequency + 
+  coef_breathing_pause * data$BreathingPause + 
+  coef_overly_sleep_frequency * data$OverlySleepy
+
+# Create a table to display
+person_stats <- data.frame(
+  Statistic = c("Weekday Sleep Duration", "Weekend Sleep Duration", "Snoring Frequency", 
+                "Breathing Pause Frequency", "Overly Sleep Frequency", "Logit(p)"),
+  Jack = c("5 hours", "8 hours", "3 (Frequently)", "3 (Frequently)", "4 (Always)", sprintf("%.3f", data$logit_p[data$Name == "Jack"])),
+  Bob = c("8 hours", "10 hours", "0 (Never)", "0 (Never)", "1 (Rarely)", sprintf("%.3f", data$logit_p[data$Name == "Bob"]))
+)
+
+# Create a nicely formatted table
+kable_output <- kable(person_stats, col.names = c("Statistic", "Jack", "Bob"), align = 'c') %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center")
+
+# Print the summary table
+kable_output
+```
+
+Interpretation of the Results From @tbl-sample-prediction
+
+**Logit(p)**: The computed logit(p) values reflect the aggregated effect of these variables.
+A positive value indicates a higher log-odds of reporting sleep troubles, while a negative value suggests lower odds.
+Jack’s logit(p) value is positive, indicating a higher likelihood of reporting sleep troubles, possibly due to his poor sleep habits and symptoms indicative of sleep disorders.
+Conversely, Bob’s negative logit(p) value suggests a lower probability of reporting sleep troubles, aligning with his generally healthier sleep patterns.
+
+**Summary**: 
+This analysis highlights how individual differences in sleep patterns and related symptoms can predict the likelihood of reporting sleep troubles.
+Jack's profile, characterized by frequent snoring, frequent breathing pauses, always feeling overly sleepy, and shorter sleep durations, aligns with a higher propensity for sleep disturbances.
+In contrast, Bob’s healthier sleep habits and fewer symptoms suggest a lower risk of sleep troubles.
+This model effectively uses sleep-related metrics to provide insights into potential sleep health outcomes.
+
+## Analysis of Residuals {#sec-residuals}
+
+In this section, we evaluate the residuals from our logistic regression model to assess the model's goodness of fit and identify any potential anomalies in our data.
+
+```{r residual-plots, echo=FALSE}
+predicted_probs <- predict(bayesian_logistic_model, type = "response")
+residuals <- residuals(bayesian_logistic_model, type = "pearson")
+
+# Plotting residuals
+plot(predicted_probs, residuals, xlab = "Predicted Probabilities", ylab = "Standardized Residuals")
+abline(h = 0, col = "red")
+```
+## Model Performance
+
 # Discussion {#sec-discussion}
 
-# Appendix {#sec-appendix}
+# Appendix {.unnumbered}
 
-## Additional Data Details {#sec-add-data}
+## A Additional Data Details {.unnumbered}
 
-## Model Details
+### A.1 Data Cleaning Process {.unnumbered}
 
-### Posterior Predictive Check {#sec-posterior}
+The data cleaning procedure for the NHANES sleep disorder data was meticulous to ensure the reliability of the dataset for analysis.
+Initially, the raw data was imported using the `read_xpt` function, ensuring that all variables of interest were accurately read into R.
+The cleaning process included the following steps:
+
+- Renaming variables for clarity.
+- Removing entries with missing values or responses that indicated uncertainty (e.g., 'don't know' or 'refused').
+- Selecting only relevant columns for the analysis.
+- Recoding the `ReportedSleepTrouble` (outcome) variable to a binary format (0 and 1).
+- Ensuring all identifiers were integers and that key sleep-related variables fell within reasonable and expected ranges.
+
+The final dataset was saved in both CSV and parquet formats for accessibility and efficient analysis.
+
+### A.2 Data Integrity and Validation {.unnumbered}
+
+Following the cleaning phase, extensive data integrity tests were conducted to validate the quality and consistency of the cleaned data.
+The `testthat` package in R was employed to programmatically assert the following:
+
+- Valid integer identifiers for each respondent.
+- Logical ranges for sleep durations on both weekdays and weekends (2 to 14 hours).
+- Appropriate categorization of snoring and breathing pause frequencies (0 to 3, indicating never to frequently).
+- Binary coding of reported sleep trouble (0 for no trouble, 1 for trouble).
+- Logical categorization of overly sleep frequency (0 to 4, indicating never to almost always).
+
+These tests are crucial for guaranteeing the analytical soundness of the dataset and were all passed successfully, adding a layer of assurance about the dataset's readiness for subsequent statistical modeling.
+
+## B Model Details {.unnumbered}
+
+### B.1 Posterior Predictive Check {.unnumbered}
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+# Set the seed for reproducibility
+set.seed(2024)
+
+# Perform posterior predictive check
+pp_check_plot <- pp_check(bayesian_logistic_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+# Plot posterior versus prior
+coef_names <- c("(Intercept)", "WeekdaySleepDuration", "WeekendSleepDuration",
+                "SnoringFrequency", "BreathingPauseFrequency", "OverlySleepFrequency")
+
+posterior_vs_prior_plot <- mcmc_areas(
+  as.array(bayesian_logistic_model),
+  pars = coef_names
+) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+# Plotting the two figures side by side using R Markdown layout options
+pp_check_plot
+posterior_vs_prior_plot
+```
 

### 2024-04-17 paper/paper.qmd
added simulation

@@
 Using logistic regression analysis on data from the NHANES 2017-March 2020 Sleep Disorders dataset, we examine how various factors such as sleep duration on weekdays and weekends, frequency of snoring, and daytime sleepiness are associated with the likelihood of reporting sleep problems to a doctor.
-The findings aim to shed light on the predictive value of self-monitored sleep behaviors in identifying individuals who may require medical attention for sleep-related issues.
+The results revealed that certain behaviors such as excessive snoring and high levels of daytime sleepiness showed a positive correlation with sleep troubles, Moreover, the duration of sleep showed a contrasting pattern of association for sleep during weekdays and weekends.
 This analysis contributes to the broader understanding of sleep health and its complex interactions with daily functioning."
@@
 The sleep disorders dataset within NHANES is enriched by questions adapted from the Munich ChronoType Questionnaire [@Roenneberg2003], targeting various aspects of sleep behavior and disorders.
-The inclusion of these questions is instrumental in exploring the complex dynamics of sleep patterns among the U.S.
-population.
+The inclusion of these questions is instrumental in exploring the complex dynamics of sleep patterns among the U.S. population.
 Due to disruptions caused by the COVID-19 pandemic, the 2019-2020 data collection cycle was prematurely halted in March 2020, leading to its combination with the 2017-2018 cycle to ensure national representativeness and analytical robustness.
@@
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 The logistic regression model is defined using a Bayesian framework, implemented in R with the `rstanarm` package.
@@
 The priors for the intercept ($\beta_0$) and the slopes ($\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$) are deliberately chosen to reflect initial neutrality or skepticism regarding the impact of each predictor on the likelihood of reporting sleep troubles.
-Specifically, $\beta_0$ is assigned a Normal(0, 1) prior, indicating our initial uncertainty about the base rate of sleep trouble in the population.
 
-For $\beta_1$ and $\beta_2$, which correspond to the effects of weekday and weekend sleep duration, we posit Normal(0, 0.5) priors.
-These priors are centered around zero, reflecting the number of sleep hour can be either positively or negatively contribute to sleep troubles, with a modest expectation of effect size.
+Specifically, the intercept $\beta_0$ and the coefficients for weekday and weekend sleep duration, $\beta_1$ and $\beta_2$, are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data.
+This reflects an initial uncertainty about the base rate of sleep trouble in the population and acknowledges that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.
 
-The coefficients $\beta_3$, $\beta_4$, $\beta_5$ relating to snoring frequency, breathing pause frequency, and overly sleep frequency are each given a Normal(0, 1) prior.
-These priors imply that, prior to analyzing the data, we do not have strong expectations about the direction or scale of these effects.
+For the predictors related to snoring frequency, breathing pause frequency, and overly sleep frequency—$\beta_3$, $\beta_4$, $\beta_5$—the same Normal(0, 1) priors are utilized, which suggests that there are no preconceived notions regarding the magnitude or direction of their impact on reporting sleep troubles.
+This choice of priors allows the data to guide the inference about these predictors without the influence of strong initial assumptions.
 
@@
 
-**Weekday Sleep Duration ($\beta_1$)**: We initially hypothesize that there may be a complex relationship between sleep duration on weekdays and sleep troubles.
+**Weekday Sleep Duration (**$\beta_1$): We initially hypothesize that there may be a complex relationship between sleep duration on weekdays and sleep troubles.
 Long sleep durations could indicate an attempt to compensate for poor quality sleep, potentially related to sleep disturbances.
@@
 
-**Weekend Sleep Duration ($\beta_2$)**: For sleep duration on weekends, we consider the possibility of a recuperative effect, where catching up on sleep may reduce the likelihood of reporting troubles.
+**Weekend Sleep Duration (**$\beta_2$): For sleep duration on weekends, we consider the possibility of a recuperative effect, where catching up on sleep may reduce the likelihood of reporting troubles.
 This hypothesis is tentative, hence a prior centered around zero with a conservative standard deviation, permitting data-driven insights into this relationship.
 
-**Snoring Frequency ($\beta_3$)**: Frequent snoring is a recognized symptom of sleep disorders like obstructive sleep apnea, which can lead to disrupted sleep.
+**Snoring Frequency (**$\beta_3$): Frequent snoring is a recognized symptom of sleep disorders like obstructive sleep apnea, which can lead to disrupted sleep.
 We posit a positive association between snoring frequency and reported sleep troubles, reflected in the priors for this coefficient.
 
-**Breathing Pause Frequency ($\beta_4$)**: Breathing pauses during sleep are indicative of sleep apnea, which is directly related to sleep troubles.
+**Breathing Pause Frequency (**$\beta_4$): Breathing pauses during sleep are indicative of sleep apnea, which is directly related to sleep troubles.
 We hypothesize a positive association for this coefficient, which is incorporated into the prior setting.
 
-**Daytime Sleepiness ($\beta_5$)**: Excessive daytime sleepiness is often the result of inadequate or disrupted nighttime sleep.
+**Daytime Sleepiness (**$\beta_5$): Excessive daytime sleepiness is often the result of inadequate or disrupted nighttime sleep.
 We hypothesize a positive relationship between daytime sleepiness and the reporting of sleep troubles, expecting that those experiencing higher levels of sleepiness will be more likely to report troubles.
@@
 ```
+
 @tbl-logistic-coefficients presents the key coefficients for understanding the influence of various sleep-related variables on the likelihood of reporting sleep troubles:
 
-**Intercept ($\beta_0$)**:
-The intercept of the model, $\beta_0$, is estimated at -1.396.
+**Intercept (**$\beta_0$): The intercept of the model, $\beta_0$, is estimated at -1.396.
 This value represents the log odds of reporting sleep troubles when all the predictors are held at zero.
@@
 
-**Weekday Sleep Duration ($\beta_1$)**:
-For the weekday sleep duration, the coefficient $\beta_1$ is 0.051, suggesting a slight increase in the log odds of reporting sleep troubles for each additional hour of sleep during the weekdays.
+**Weekday Sleep Duration (**$\beta_1$): For the weekday sleep duration, the coefficient $\beta_1$ is 0.051, suggesting a slight increase in the log odds of reporting sleep troubles for each additional hour of sleep during the weekdays.
 This could be indicative of a situation where those who sleep more on weekdays might be doing so due to sleep issues that have led them to consult with a doctor.
 
-**Weekend Sleep Duration ($\beta_2$)**:
-Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131.
+**Weekend Sleep Duration (**$\beta_2$): Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131.
 This implies that for each additional hour of sleep during the weekends, there is a decrease in the log odds of reporting sleep troubles.
@@
 
-**Snoring Frequency ($\beta_3$)**:
-The coefficient for snoring frequency, $\beta_3$, has a value of 0.032.
+**Snoring Frequency (**$\beta_3$): The coefficient for snoring frequency, $\beta_3$, has a value of 0.032.
 While this is a positive value, suggesting that an increase in snoring frequency is associated with an increase in the likelihood of reporting sleep troubles, the effect size is relatively small.
 
-**Breathing Pause Frequency ($\beta_4$)**:
-The coefficient for breathing pause frequency, $\beta_4$, is 0.412, indicating a more substantial positive association with reporting sleep troubles.
+**Breathing Pause Frequency (**$\beta_4$): The coefficient for breathing pause frequency, $\beta_4$, is 0.412, indicating a more substantial positive association with reporting sleep troubles.
 This aligns with clinical understanding, as breathing pauses are often associated with sleep disorders like sleep apnea, which can be a significant concern prompting medical consultation.
 
-**Overly Sleep Frequency ($\beta_5$)**:
-Lastly, the coefficient for overly sleep frequency, $\beta_5$, is 0.425, suggesting a strong relationship with the reporting of sleep troubles.
+**Overly Sleep Frequency (**$\beta_5$): Lastly, the coefficient for overly sleep frequency, $\beta_5$, is 0.425, suggesting a strong relationship with the reporting of sleep troubles.
 This result is intuitive, as feeling excessively sleepy can be a direct symptom of poor sleep quality or a sleep disorder, leading to discussions with a healthcare provider.
 
-**Interpretation**:
-The model indicates that various factors are related to the reporting of sleep troubles.
+**Interpretation**: The model indicates that various factors are related to the reporting of sleep troubles.
 While longer weekday sleep duration slightly increases the log odds of reporting sleep troubles, longer weekend sleep can reduce it.
@@
 \end{align}
-
+```
 ## Predict Sleep Disorder Through Self Reported Patterns
@@
 
-**Summary**: 
-This analysis highlights how individual differences in sleep patterns and related symptoms can predict the likelihood of reporting sleep troubles.
+**Summary**: This analysis highlights how individual differences in sleep patterns and related symptoms can predict the likelihood of reporting sleep troubles.
 Jack's profile, characterized by frequent snoring, frequent breathing pauses, always feeling overly sleepy, and shorter sleep durations, aligns with a higher propensity for sleep disturbances.
@@
 
-## Analysis of Residuals {#sec-residuals}
+## Model Performance
 
-In this section, we evaluate the residuals from our logistic regression model to assess the model's goodness of fit and identify any potential anomalies in our data.
+```{r}
+#| echo: false
+# Predict probabilities on the training data
+predicted_probabilities <- predict(bayesian_logistic_model, type = "response")
 
-```{r residual-plots, echo=FALSE}
-predicted_probs <- predict(bayesian_logistic_model, type = "response")
-residuals <- residuals(bayesian_logistic_model, type = "pearson")
+# Set a new threshold for classification
+threshold <- 0.5  # Adjust this value as needed
 
-# Plotting residuals
-plot(predicted_probs, residuals, xlab = "Predicted Probabilities", ylab = "Standardized Residuals")
-abline(h = 0, col = "red")
+# Calculate the predicted classes based on the threshold
+predicted_classes <- ifelse(predicted_probabilities > threshold, 1, 0)
+
+# Actual classes from the data
+actual_classes <- cleaned_data$ReportedSleepTrouble
+
+# Create a confusion matrix
+conf_matrix <- table(Predicted = predicted_classes, Actual = actual_classes)
+
+# Extract components and round them to integer
+TP <- as.integer(conf_matrix["1", "1"])  # True Positives
+TN <- as.integer(conf_matrix["0", "0"])  # True Negatives
+FP <- as.integer(conf_matrix["1", "0"])  # False Positives
+FN <- as.integer(conf_matrix["0", "1"])  # False Negatives
+
+# Calculate accuracy and round rates to three decimal places
+accuracy <- round(sum(diag(conf_matrix)) / sum(conf_matrix), 3)
+TP_rate <- round(TP / (TP + FN), 3)
+TN_rate <- round(TN / (TN + FP), 3)
+FP_rate <- round(FP / (FP + TN), 3)
+FN_rate <- round(FN / (FN + TP), 3)
+
+# Define metrics data frames
+counts_df <- data.frame(
+  Metric = c("True Positives", "True Negatives", "False Positives", "False Negatives"),
+  Count = c(TP, TN, FP, FN)
+)
+
+rates_df <- data.frame(
+  Metric = c("Accuracy", "True Positive Rate", "True Negative Rate", "False Positive Rate", "False Negative Rate"),
+  Rate = c(accuracy, TP_rate, TN_rate, FP_rate, FN_rate)
+)
+
+# Create tables with kable and apply styling
+counts_table <- kable(counts_df, col.names = c("Metric", "Count"), align = c('l', 'r')) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
+
+rates_table <- kable(rates_df, col.names = c("Metric", "Rate"), align = c('l', 'r')) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
 ```
-## Model Performance
+
+```{r}
+#| label: tbl-performance-count
+#| tbl-cap: "Count Metrics of the Model Performance"
+#| echo: false
+# Generate the table for counts
+counts_table <- kable(counts_df, 
+                       col.names = c("Metric", "Count"), 
+                       align = c('l', 'r')) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
+
+#Print the table
+counts_table
+```
+@tbl-performance-count showcases how well the model predicts reported sleep troubles with counts of true positives and true negatives indicating correct predictions for the presence and absence of sleep trouble, respectively.
+Conversely, false positives and false negatives highlight where the model predicts incorrectly, either by overestimating (predicting sleep trouble where there is none) or underestimating (failing to detect reported sleep trouble) the condition.
+These counts are essential for assessing the model's performance accuracy, balancing the sensitivity to detect actual cases against the specificity to dismiss non-cases.
+
+```{r}
+#| label: tbl-performance-rate
+#| tbl-cap: "Rate Metrics of the Model Performance"
+#| echo: false
+# Generate the table for rates
+rates_table <- kable(rates_df, 
+                     col.names = c("Metric", "Rate"), 
+                     align = c('l', 'r')) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
+
+#Print the table
+rates_table
+```
+
+@tbl-performance-rate shows the performance indicators of the model.
+The accuracy stands at 75.3%, denoting the proportion of predictions that the model got right.
+The true positive rate is 15.3%, reflecting the model’s efficiency in correctly predicting actual instances of sleep trouble.
+Impressively, the true negative rate is 96.4%, indicating the model's strong capability to recognize individuals without sleep trouble.
+Conversely, the false positive rate is minimal at 3.6%, suggesting that the model seldom incorrectly flags sleep trouble where there is none.
+However, the model’s potential area for improvement is highlighted by the false negative rate of 84.7%, which points to a substantial number of actual sleep trouble cases that the model fails to detect.
+These metrics are instrumental for evaluating the model's precision and guiding further refinements.
 
@@
 
+## Advantages of Self-Reported Data in Sleep Health
+
+Using self-reported patterns to predict sleep troubles offers a valuable, accessible means for identifying individuals who may require medical intervention.
+This approach leverages personal sleep logs and questionnaires, which are cost-effective and can be implemented on a large scale, enhancing the reach of health monitoring.
+Self-reported data provide insights into an individual's sleep duration, frequency of disturbances like snoring and breathing pauses, and subjective feelings of sleepiness, all of which are crucial for diagnosing sleep disorders.
+
+## Impact of Sleep Patterns on Reported Sleep Troubles {#sec-sleep-pattern-impact}
+
+The analysis presented in this paper underscores the complex interplay between various sleep behaviors and the likelihood of individuals reporting sleep troubles.
+Our model's findings align with existing literature suggesting that both insufficient sleep duration and poor sleep quality are significant predictors of sleep disorders [@State_2024].
+In particular, the rates of false negatives indicate a substantial portion of individuals experiencing sleep issues that the model failed to capture, possibly reflecting the nuanced and subjective nature of sleep disturbances.
+
+## The Role of Breathing Irregularities and Daytime Sleepiness {#sec-breathing-daytime-sleep}
+
+Breathing irregularities and excessive daytime sleepiness emerged as strong predictors in our model, pointing towards conditions such as sleep apnea, which remains underdiagnosed in the general population [@Foundation_2021].
+The high false negative rate may partly result from such conditions' subtler signs, which individuals may not report or may attribute to other causes.
+
+## Assessing Model Performance and Clinical Implications {#sec-model-performance}
+
+While the model demonstrated robust true negative rates, indicating a strong ability to correctly identify individuals without sleep troubles, the high false negative rate poses significant concerns.
+Clinically, this could lead to missed diagnoses and a delay in treatment for those with unreported sleep disorders.
+Future iterations of the model might benefit from incorporating additional predictors or employing different statistical techniques to enhance sensitivity.
+
+## Limitations and Future Directions {#sec-limitations-future-directions}
+
+### Limitations {#sec-limitations}
+
+A significant limitation of this analysis was the sole reliance on self-reported data, which can introduce response bias and affect the accuracy of the findings.
+Additionally, the dataset may not adequately represent the entire population, as some demographics may be underrepresented.
+An inherent challenge also lay in the absence of a validation set, which could lead to an overfit model that doesn't generalize well to new, unseen data.
+
+### Future Directions {#sec-future-directions}
+
+For future research, it would be beneficial to conduct longitudinal studies to understand the temporal aspects of sleep patterns better and their long-term health impacts.
+Utilizing a separate validation dataset would help in assessing the model's predictive power and in refining the classification threshold to balance the trade-off between sensitivity and specificity effectively.
+Incorporating objective data sources, such as wearable technology that provides sleep quality indicators, could greatly enhance the robustness of the findings.
+Additionally, employing advanced machine learning algorithms might reveal complex non-linear interactions and patterns within the predictors, offering richer insights than those obtainable from traditional logistic regression analysis.
+Emphasizing the development of predictive models that are both interpretable and accurate will be key to advancing our understanding of sleep health.
+
+\newpage
 # Appendix {.unnumbered}
@@
 
+From @fig-ppcheckandposteriorvsprior, from the left plot the alignment of the observed data (solid black line) with the simulations from the model (lighter blue lines) suggests that the model captures the essential trends and variability in the data.
+This congruence indicates a model that adequately reflects the observed phenomena and provides assurance that the predictive capacity of the model is robust for the data in question.
+
+The right plot displays the comparison between the posterior and prior distributions which offers insights into the influence of the data on our initial assumptions.
+The posterior distributions remain close to the priors, suggesting that the priors were well-calibrated and that the data has refined, rather than overturned, these initial beliefs.
+The spread of the posterior distributions signifies the inherent uncertainty and variability within the model, which is a fundamental aspect of statistical modeling and Bayesian analysis.
+
+\newpage
+
 # References

### 2024-04-17 scripts/00_simulate-data.R

@@ -1,0 +1,51 @@
+#### Preamble ####
+# Purpose: To simulate a dataset resembling NHANES sleep disorder data
+# Author: Terry Tu
+# Date: 17 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: tidyverse package
+
+# Load necessary libraries
+library(tidyverse)
+
+# Set a seed for reproducibility
+set.seed(2024)
+
+# Define the number of individuals to simulate
+n_people <- 1000
+
+# Simulate weekday and weekend sleep duration based on normal distributions
+weekday_sleep <- rnorm(n_people, mean = 7, sd = 1.5)
+weekend_sleep <- rnorm(n_people, mean = 8, sd = 2)
+
+# Simulate snoring frequency based on a categorical distribution
+snoring_frequency <- sample(0:3, n_people, replace = TRUE, prob = c(0.5, 0.2, 0.2, 0.1))
+
+# Simulate breathing pause frequency assuming it's less common
+breathing_pause_frequency <- sample(0:3, n_people, replace = TRUE, prob = c(0.7, 0.15, 0.1, 0.05))
+
+# Simulate reported sleep trouble based on a binomial distribution where probability(p) depends on other factors
+p_sleep_trouble <- plogis(-0.5 + 0.1 * weekday_sleep - 0.05 * weekend_sleep + 0.2 * snoring_frequency + 0.3 * breathing_pause_frequency)
+reported_sleep_trouble <- rbinom(n_people, 1, p_sleep_trouble)
+
+# Combine all variables into a data frame
+simulated_sleep_data <- tibble(
+  WeekdaySleepDuration = weekday_sleep,
+  WeekendSleepDuration = weekend_sleep,
+  SnoringFrequency = snoring_frequency,
+  BreathingPauseFrequency = breathing_pause_frequency,
+  ReportedSleepTrouble = reported_sleep_trouble
+)
+
+# Inspect the first few rows of the simulated data
+head(simulated_sleep_data)
+
+# Plot the relationship between sleep duration and reported trouble
+ggplot(simulated_sleep_data, aes(x = WeekdaySleepDuration, fill = as.factor(ReportedSleepTrouble))) +
+  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
+  labs(title = "Simulated Data: Weekday Sleep Duration and Reported Sleep Trouble",
[…]

### 2024-04-18 paper/paper.qmd
added download script

@@
 Impressively, the true negative rate is 96.4%, indicating the model's strong capability to recognize individuals without sleep trouble.
-Conversely, the false positive rate is minimal at 3.6%, suggesting that the model seldom incorrectly flags sleep trouble where there is none.
-However, the model’s potential area for improvement is highlighted by the false negative rate of 84.7%, which points to a substantial number of actual sleep trouble cases that the model fails to detect.
+Conversely, the false positive (type I error) rate is minimal at 3.6%, suggesting that the model seldom incorrectly flags sleep trouble where there is none.
+However, the model’s potential area for improvement is highlighted by the false negative (type II error) rate of 84.7%, which points to a substantial number of actual sleep trouble cases that the model fails to detect.
 These metrics are instrumental for evaluating the model's precision and guiding further refinements.

### 2024-04-18 scripts/01_download-data.R

@@ -1,0 +1,24 @@
+#### Preamble ####
+# Purpose: To download the 2017-2020 NHANES Sleep Disorders Data
+# Author: Terry Tu
+# Date: 17 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'here' package, Internet connection required
+
+# Load the here package
+library(here)
+
+# Define the URL of the NHANES Sleep Disorders Data
+url <- "https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/P_SLQ.XPT"
+
+# Use the here function to create a file path
+destination <- here("data", "raw_data", "P_SLQ.XPT")
+
+# Create the directory if it doesn't exist
+if (!dir.exists(dirname(destination))) {
+  dir.create(dirname(destination), recursive = TRUE)
+}
+
+# Download the file
+download.file(url, destfile = destination, mode = "wb")

### 2024-04-18 paper/paper.qmd
added setup script

@@
 title-block-banner: true
-abstract: "This study investigates the potential relationship between individuals' sleep patterns and the self-reported incidence of sleep troubles as confirmed by a medical professional.
-Using logistic regression analysis on data from the NHANES 2017-March 2020 Sleep Disorders dataset, we examine how various factors such as sleep duration on weekdays and weekends, frequency of snoring, and daytime sleepiness are associated with the likelihood of reporting sleep problems to a doctor.
-The results revealed that certain behaviors such as excessive snoring and high levels of daytime sleepiness showed a positive correlation with sleep troubles, Moreover, the duration of sleep showed a contrasting pattern of association for sleep during weekdays and weekends.
-This analysis contributes to the broader understanding of sleep health and its complex interactions with daily functioning."
+abstract: "This study uses logistic regression to analyze the relationship between self-reported sleep behaviors and the likelihood of reporting sleep troubles, using data from the NHANES 2017 to March 2020 Sleep Disorders dataset.
+We examine factors such as snoring frequency, daytime sleepiness, and sleep duration on weekdays and weekends.
+Findings indicate that excessive snoring and high daytime sleepiness are strongly associated with more reported sleep issues, while longer weekend sleep correlates with fewer reports.
+These results highlight the importance of good sleep hygiene for overall health and provide insights into how daily behaviors affect sleep quality."
 thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Sleep-Disorder-Analysis"
@@
 
-The rhythm of our nightly rest is more than a personal habit; it's a public health pulse that captures the essence of well-being in our fast-paced society.
-The increasing prevalence of sleep disorders and their impact on daily life and overall health has become a point of societal concern, akin to the growing conversation around mental health and lifestyle diseases.
-The intimate link between quality of sleep and the vitality of an individual's life prompts a closer examination of sleep patterns within the populace.
-Drawing on the rich dataset from the National Health and Nutrition Examination Survey (NHANES) for the years 2017 to March 2020, our research delves into the self-reported instances of sleep disturbances and their association with various sleep behaviors.
+Sleep disturbances are increasingly recognized as a major public health issue, comparable to concerns surrounding mental health and chronic lifestyle diseases.
+The effects of sleep disorders extend beyond personal inconvenience, influencing public health, workplace productivity, and overall quality of life.
+Despite increased awareness, there is still a significant gap in understanding the behaviors and patterns linked to sleep disturbances.
+This study addresses this gap by analyzing self-reported data on sleep behaviors and their impact on health outcomes, and by developing tools for individuals to predict potential sleep issues.
 
-This study is an explorative journey into the silent epidemic of sleep disorders that plague modern society, affecting productivity, mental health, and long-term wellbeing.
-The aim is to uncover the underlying patterns of sleep behavior that correlate with the reports of sleep troubles to medical professionals, thereby piecing together the nocturnal puzzle of restless societies.
-We explore the quantitative relationship between self-reported snoring frequency, feelings of daytime sleepiness, and the regularity of sleep hours during weekdays and weekends with the likelihood of reporting sleep issues to a doctor [@Prevention_2021].
+Utilizing the National Health and Nutrition Examination Survey (NHANES) dataset from 2017 to March 2020, our research examines how various sleep behaviors—such as snoring frequency, daytime sleepiness, and sleep regularity—correlate with the likelihood of reporting sleep issues to healthcare providers.
+The primary estimand of our analysis is the association between these behaviors and the probability of an individual reporting sleep disturbances, quantified through logistic regression.
+This approach provides a clearer understanding of the epidemiology of sleep problems and identifies key predictive factors.
 
-Our analysis hinges on the application of logistic regression to the NHANES dataset, which presents a comprehensive view of American sleep habits.
-By interpreting the nuances of this rich dataset, we aspire to illuminate the factors that signal the need for medical attention in the domain of sleep health.
-The outcome of our investigation is poised to provide a scaffold for healthcare professionals and policymakers to base early intervention strategies, aiming to cultivate a well-rested population.
+Our findings pinpoint specific behaviors that significantly correlate with sleep disturbances.
+These insights are vital for healthcare professionals and policy makers, identifying potential areas for intervention to improve sleep health.
+The impact of our research is substantial, providing a basis for strategies to reduce the effects of sleep disorders on individuals and on society as a whole.
 
-Following this introduction, the structure of the paper is laid out to facilitate a coherent flow of information and analysis.
-@sec-data (Data) provides a meticulous breakdown of the NHANES dataset, elucidating the data cleaning process and offering a descriptive overview of the key variables.
-@sec-model (Model) details the logistic regression model's design and the rationale behind the choice of predictors.
-@sec-result (Result) presents the findings, interpreted with precision and caution, alongside graphical representations for clarity.
-Concluding the paper, @sec-discussion (Discussion) reflects on the broader implications of the study, acknowledging limitations and proposing avenues for future research.
+By establishing a direct link between certain sleep behaviors and reported sleep issues, our study contributes to ongoing efforts to address the silent epidemic of sleep disorders that affects a large segment of the population.
+It also offers individuals a valuable tool to assess their risk of sleep disturbances, promoting early detection and proactive health management.
+
+The paper is structured to ensure a clear presentation of our research and findings:
+@sec-data (Data) describes our data collection and cleaning processes, providing a detailed overview of the NHANES dataset.
+@sec-model (Model) explains the logistic regression model, including the selection of predictors and the underlying rationale.
+@sec-result (Result) presents our main findings, supported by graphical representations for clearer understanding.
+Concluding the paper, @sec-discussion (Discussion) provides the implications of our research, recognizing limitations and suggesting future research directions.
 
@@
 ```
+
 The logistic regression model is defined using a Bayesian framework, implemented in R with the `rstanarm` package.
-This approach allows us to incorporate prior knowledge about the parameters and to estimate their posterior distributions based on the observed data.
+This approach enables the integration of prior knowledge about the parameters to estimate their posterior distributions based on the observed data.
 
@@
 The log odds of reporting sleep troubles are modeled as a linear combination of the predictors, with $\beta_0$ representing the intercept, and $\beta_1$ through $\beta_5$ representing the slopes for the respective predictors.
+Among these, $\beta_1$ (weekday sleep duration) and $\beta_2$ (weekend sleep duration) are continuous variables, reflecting the number of hours slept.
+In contrast, $\beta_3$ (snoring frequency), $\beta_4$ (breathing pause frequency), and $\beta_5$ (overly sleep frequency) are treated as ordinal categorical variables with numerically encoded categories.
 
-The priors for the intercept ($\beta_0$) and the slopes ($\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$) are deliberately chosen to reflect initial neutrality or skepticism regarding the impact of each predictor on the likelihood of reporting sleep troubles.
+The priors for the intercept ($\beta_0$) and the slopes are deliberately chosen to reflect initial neutrality or skepticism about each predictor's impact on the likelihood of reporting sleep troubles.
+Specifically, $\beta_0$, $\beta_1$, and $\beta_2$ are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data, acknowledging that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.
 
-Specifically, the intercept $\beta_0$ and the coefficients for weekday and weekend sleep duration, $\beta_1$ and $\beta_2$, are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data.
-This reflects an initial uncertainty about the base rate of sleep trouble in the population and acknowledges that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.
-
-For the predictors related to snoring frequency, breathing pause frequency, and overly sleep frequency—$\beta_3$, $\beta_4$, $\beta_5$—the same Normal(0, 1) priors are utilized, which suggests that there are no preconceived notions regarding the magnitude or direction of their impact on reporting sleep troubles.
-This choice of priors allows the data to guide the inference about these predictors without the influence of strong initial assumptions.
+For $\beta_3$, $\beta_4$, and $\beta_5$, which relate to snoring frequency, breathing pause frequency, and overly sleep frequency, the same Normal(0, 1) priors are used.
+This choice indicates no strong preconceived notions about the direction or magnitude of their impact, allowing the data itself to guide the inference about these predictors without the influence of strong initial assumptions.
 
@@
 
-**Intercept (**$\beta_0$): The intercept of the model, $\beta_0$, is estimated at -1.396.
-This value represents the log odds of reporting sleep troubles when all the predictors are held at zero.
-It suggests that, in the absence of any sleep issues or without considering any sleep behaviors, the baseline log odds of reporting sleep troubles to a doctor are negative, indicating a lower probability of reporting sleep troubles.
+**Intercept ($\beta_0$)**: The intercept of the model, $\beta_0$, is estimated at -1.396.
+This value represents the log odds of reporting sleep troubles when all predictors are at their reference levels (minimal values).
+It suggests a lower likelihood of reporting sleep troubles in the absence of any problematic sleep behaviors.
 
-**Weekday Sleep Duration (**$\beta_1$): For the weekday sleep duration, the coefficient $\beta_1$ is 0.051, suggesting a slight increase in the log odds of reporting sleep troubles for each additional hour of sleep during the weekdays.
-This could be indicative of a situation where those who sleep more on weekdays might be doing so due to sleep issues that have led them to consult with a doctor.
+**Weekday Sleep Duration ($\beta_1$)**: The coefficient $\beta_1$ is 0.051 for weekday sleep duration, indicating that each additional hour of sleep on weekdays slightly increases the log odds of reporting sleep troubles.
+This might suggest that individuals sleeping more on weekdays could be experiencing underlying sleep issues that lead them to consult a doctor.
 
-**Weekend Sleep Duration (**$\beta_2$): Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131.
-This implies that for each additional hour of sleep during the weekends, there is a decrease in the log odds of reporting sleep troubles.
-It could be interpreted as those who manage to sleep more on weekends are less likely to report sleep problems, possibly because catching up on sleep may alleviate some of their weekday sleep deficits.
+**Weekend Sleep Duration ($\beta_2$)**: Conversely, the coefficient for weekend sleep duration, $\beta_2$, is -0.131.
+This implies that each additional hour of sleep during weekends decreases the log odds of reporting sleep troubles, supporting the idea that extra sleep on weekends might help reduce the chance of reporting sleep trouble.
 
-**Snoring Frequency (**$\beta_3$): The coefficient for snoring frequency, $\beta_3$, has a value of 0.032.
-While this is a positive value, suggesting that an increase in snoring frequency is associated with an increase in the likelihood of reporting sleep troubles, the effect size is relatively small.
+**Snoring Frequency ($\beta_3$)**: The coefficient of 0.032 for snoring frequency indicates a positive but small effect on the likelihood of reporting sleep troubles.
+This relationship suggests that as snoring frequency increases, so does the likelihood of reporting sleep troubles, although the impact is relatively modest compared to other factors.
 
-**Breathing Pause Frequency (**$\beta_4$): The coefficient for breathing pause frequency, $\beta_4$, is 0.412, indicating a more substantial positive association with reporting sleep troubles.
-This aligns with clinical understanding, as breathing pauses are often associated with sleep disorders like sleep apnea, which can be a significant concern prompting medical consultation.
+**Breathing Pause Frequency ($\beta_4$)**: A more substantial coefficient of 0.412 for breathing pause frequency reflects a significant increase in the likelihood of reporting sleep troubles.
+Breathing pauses are commonly associated with conditions like sleep apnea, which are often serious enough to warrant medical attention.
 
-**Overly Sleep Frequency (**$\beta_5$): Lastly, the coefficient for overly sleep frequency, $\beta_5$, is 0.425, suggesting a strong relationship with the reporting of sleep troubles.
-This result is intuitive, as feeling excessively sleepy can be a direct symptom of poor sleep quality or a sleep disorder, leading to discussions with a healthcare provider.
+**Overly Sleep Frequency ($\beta_5$)**: Similarly, the coefficient for overly sleep frequency is 0.425, indicating a strong positive association with reporting sleep troubles.
+Frequent excessive daytime sleepiness is a critical indicator of poor sleep quality or underlying sleep disorders.
 
-**Interpretation**: The model indicates that various factors are related to the reporting of sleep troubles.
-While longer weekday sleep duration slightly increases the log odds of reporting sleep troubles, longer weekend sleep can reduce it.
-Frequent snoring has a smaller effect compared to breathing pauses and daytime sleepiness, which both have significant positive associations with reporting sleep troubles.
-These findings underline the complexity of sleep behavior and its impact on sleep quality and health consultations.
+**Summary**: This analysis highlights the differential impacts of various sleep behaviors on the likelihood of reporting sleep troubles.
+Notably, while longer sleep on weekends appears protective, increased weekday sleep duration and higher frequencies of snoring, breathing pauses, and daytime sleepiness are linked with an elevated likelihood of reporting sleep troubles.
+These insights could guide clinical assessments and interventions aimed at improving sleep health.
 
@@
 \begin{align}
-\text{logit}(p_i) &= -1.396 + 0.051 \times x_{i1} - 0.131 \times x_{i2} + 0.032 \times x_{i3} + 0.412 \times x_{i4} + 0.425 \times x_{i5} + \varepsilon_i
+\text{logit}(\hat{p_i}) &= -1.396 + 0.051 \times x_{i1} - 0.131 \times x_{i2} + 0.032 \times x_{i3} + 0.412 \times x_{i4} + 0.425 \times x_{i5} \\
+\hat{p_i} &= \frac{1}{1+e^{-logit(\hat{p_i})}}
 \end{align}
 ```
+
+The above equation represents the logistic regression model used to estimate the probability ($\hat{p_i}$) of an individual reporting sleep troubles.
+In this model, $\text{logit}(\hat{p_i})$ is the log-odds of the individual reporting sleep troubles, which is calculated as a linear combination of several predictors.
+Here, $\beta_0=−1.396$ is the intercept, which represents the log-odds of reporting sleep troubles when all predictors are at zero.
+The coefficients $\beta_1=0.051$, $\beta_2=−0.131$, $\beta_3=0.032$, $\beta_4=0.412$, and $\beta_5=0.425$ represent the change in log-odds of reporting sleep troubles for each one-unit change in weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$), respectively (Please check @sec-variable for details).
+This model assumes that the probability of reporting sleep troubles increases or decreases with changes in these sleep-related behaviors and symptoms.
+
 ## Predict Sleep Disorder Through Self Reported Patterns
@@
 #| echo: false
+
 # Define the coefficients from the model
@@
 
+# Calculate the probability p from logit(p)
+data$probability_p <- exp(data$logit_p) / (1 + exp(data$logit_p))
+
 # Create a table to display
@@
   Statistic = c("Weekday Sleep Duration", "Weekend Sleep Duration", "Snoring Frequency", 
-                "Breathing Pause Frequency", "Overly Sleep Frequency", "Logit(p)"),
-  Jack = c("5 hours", "8 hours", "3 (Frequently)", "3 (Frequently)", "4 (Always)", sprintf("%.3f", data$logit_p[data$Name == "Jack"])),
-  Bob = c("8 hours", "10 hours", "0 (Never)", "0 (Never)", "1 (Rarely)", sprintf("%.3f", data$logit_p[data$Name == "Bob"]))
+                "Breathing Pause Frequency", "Overly Sleep Frequency", "Logit(p)", "Probability(p)"),
+  Jack = c("5 hours", "8 hours", "3 (Frequently)", "3 (Frequently)", "4 (Always)", sprintf("%.3f", data$logit_p[1]), sprintf("%.3f", data$probability_p[1])),
+  Bob = c("8 hours", "10 hours", "0 (Never)", "0 (Never)", "1 (Rarely)", sprintf("%.3f", data$logit_p[2]), sprintf("%.3f", data$probability_p[2]))
 )
@@
 kable_output
+
 ```
@@
 
-**Logit(p)**: The computed logit(p) values reflect the aggregated effect of these variables.
-A positive value indicates a higher log-odds of reporting sleep troubles, while a negative value suggests lower odds.
-Jack’s logit(p) value is positive, indicating a higher likelihood of reporting sleep troubles, possibly due to his poor sleep habits and symptoms indicative of sleep disorders.
-Conversely, Bob’s negative logit(p) value suggests a lower probability of reporting sleep troubles, aligning with his generally healthier sleep patterns.
+**Logit(p) and Probability(p)**: The values calculated from the logistic regression model indicate the log-odds and probability of reporting sleep troubles for each individual.
+A positive logit(p) value, such as Jack's 0.843, suggests higher log-odds of reporting sleep troubles, which translates to a probability of 69.9%.
+This indicates a substantial likelihood of sleep issues, possibly reflecting Jack's problematic sleep patterns and symptoms indicative of sleep disorders, such as frequent snoring, regular breathing pauses, and constant daytime sleepiness.
 
-**Summary**: This analysis highlights how individual differences in sleep patterns and related symptoms can predict the likelihood of reporting sleep troubles.
-Jack's profile, characterized by frequent snoring, frequent breathing pauses, always feeling overly sleepy, and shorter sleep durations, aligns with a higher propensity for sleep disturbances.
-In contrast, Bob’s healthier sleep habits and fewer symptoms suggest a lower risk of sleep troubles.
-This model effectively uses sleep-related metrics to provide insights into potential sleep health outcomes.
+Conversely, Bob’s logit(p) value of -1.873 presents negative log-odds, correlating with a much lower probability of 13.3% of reporting sleep troubles.
+This outcome aligns with his healthier sleep behaviors, such as less frequent snoring and almost no daytime sleepiness, which likely contribute to a significantly reduced risk of sleep disturbances.
+
+**Summary**: This analysis shows the influence of individual sleep behaviors and symptoms on the probability of reporting sleep troubles.
+Jack’s profile, marked by severe symptoms and less restful sleep, is predictive of a higher likelihood of experiencing sleep-related issues.
+In contrast, Bob's minimal sleep disturbances and healthier sleep habits are associated with a lower probability of sleep troubles.
+This modeling approach effectively leverages sleep-related metrics to provide insights into potential sleep health outcomes, emphasizing the predictive power of individual sleep patterns and behaviors in identifying potential sleep problems.
 
@@
 From @fig-ppcheckandposteriorvsprior, from the left plot the alignment of the observed data (solid black line) with the simulations from the model (lighter blue lines) suggests that the model captures the essential trends and variability in the data.
-This congruence indicates a model that adequately reflects the observed phenomena and provides assurance that the predictive capacity of the model is robust for the data in question.
+This alignment indicates a model that adequately reflects the observed trend and provides assurance that the predictive capacity of the model is robust for the data in question.
 

### 2024-04-18 scripts/00_environment-setup.R

@@ -1,0 +1,29 @@
+#### Preamble ####
+# Purpose: Set up the environment by installing and loading required R packages
+# Author: Terry Tu
+# Date: 17 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+
+# Define a vector of required packages
+required_packages <- c(
+  "ggplot2", "dplyr", "tidyverse", "here", "readr", 
+  "knitr", "kableExtra", "rstanarm", "haven", "testthat",
+  "arrow", "gridExtra", "modelsummary", "bayesplot"
+)
+
+# Function to check and install any missing packages
+install_and_load <- function(packages) {
+  for (package in packages) {
+    if (!require(package, character.only = TRUE)) {
+      install.packages(package)
+      library(package, character.only = TRUE)
+    }
+  }
+}
+
+# Install and load the packages
+install_and_load(required_packages)
+
+# Print a message to indicate completion
+cat("All required packages are installed and loaded.\n")

### 2024-04-18 paper/datasheet.qmd
added datasheet

---
title: "Datasheet for Decoding Sleep Disorders Through Self-Reported Patterns"
abstract: "This study uses logistic regression to analyze the relationship between self-reported sleep behaviors and the likelihood of reporting sleep troubles, using data from the NHANES 2017 to March 2020 Sleep Disorders dataset. We examine factors such as snoring frequency, daytime sleepiness, and sleep duration on weekdays and weekends. Findings indicate that excessive snoring and high daytime sleepiness are strongly associated with more reported sleep issues, while longer weekend sleep correlates with fewer reports. These results highlight the importance of good sleep hygiene for overall health and provide insights into how daily behaviors affect sleep quality."
thanks: "Code and data supporting this analysis are available at https://github.com/TEJMaster/Sleep-Disorder-Analysis"
author: "Terry Tu"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
link-citations: true
---

Extract of the questions from @gebru2021datasheets.

# Motivation

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
   - The dataset was created to examine the relationships between various sleep-related factors and reported sleep troubles. It aims to analyze the influence of sleep duration, sleep habits, and sleep disorders based on data collected prior to the COVID-19 pandemic disruptions.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
   - The dataset was compiled by the National Center for Health Statistics (NCHS) as part of the National Health and Nutrition Examination Survey (NHANES).

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
   - The dataset creation was funded by the Centers for Disease Control and Prevention (CDC), an agency within the Department of Health and Human Services (HHS) of the United States.

4. *Any other comments?*
   - The dataset includes important adaptations to ensure continuity and representativeness despite disruptions caused by the COVID-19 pandemic.

# Composition

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
   - Each instance represents data from an individual participant aged 16 or older, detailing their sleep patterns and associated disorders as part of the NHANES 2017-March 2020 cycle.

2. *How many instances are there in total (of each type, if appropriate)?*
   - The exact number of instances corresponds to the number of participants surveyed during the NHANES cycles from 2017 to March 2020. However, specific numbers can be derived from the NHANES dataset documentation.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
   - This dataset is a nationally representative sample combining two NHANES cycles due to the incomplete data collection caused by the COVID-19 pandemic. Special weighting processes were applied to ensure representativeness.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
   - Each instance includes structured data about the participant’s usual sleep duration on weekdays and weekends, sleep habits, and reported sleep disorders, among other health-related information.

5. *Is there a label or target associated with each instance? If so, please provide a description.*
   - The primary labels include sleep duration (weekdays and weekends) and the presence of sleep disorders, used to analyze patterns and correlations in sleep behavior.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
   - Some instances may have missing data due to non-response or inconsistencies during data collection, which are handled according to NHANES data processing protocols.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
   - No explicit relationships between instances are noted; each instance is treated as independent for the purposes of public health analysis.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
   - No specific data splits are recommended by NHANES; researchers may determine splits based on their analytical needs.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
   - Errors and noise might be present due to the self-reported nature of some data and the limitations in data collection methodologies, as noted in the NHANES documentation.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
    - The dataset is self-contained and does not rely on external resources for its primary use. However, additional details and guidelines can be accessed through the NHANES website.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
    - No confidential data is included; all NHANES data is anonymized and made public for research purposes under strict ethical guidelines.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
    - No, the dataset contains health-related information that is scientific and non-offensive in nature.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
    - The dataset includes demographic variables such as age, gender, and ethnicity, allowing for detailed subgroup analyses within the population.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
    - NHANES data is designed to be anonymous, with no direct or indirect identifiers included in the dataset.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
    - The dataset includes sensitive health data, which is handled according to the ethical guidelines and privacy policies mandated by the CDC and HHS.

16. *Any other comments?*
    - Researchers are advised to follow appropriate ethical guidelines when utilizing this dataset, particularly when dealing with sensitive health information.

# Collection Process

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
    - Data was directly collected through in-person interviews and examinations using the Computer-Assisted Personal Interview (CAPI) system, ensuring high data quality and reliability. The interviews were conducted by trained personnel, and the collected data underwent quality checks to validate the accuracy.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
    - The data was collected using standardized procedures and the CAPI system, which includes built-in consistency checks. Additional quality control measures included reviewing a percentage of the audio recordings of the interviews to validate data accuracy.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    - The dataset is a nationally representative sample designed to reflect the U.S. population, using stratified multistage sampling. The sample includes individuals from various demographics and geographic locations across the country.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
    - Data collection was performed by trained NHANES staff who are part of the NCHS. These staff members are trained healthcare professionals and data collectors employed by the government, and compensation is part of their regular salaries.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
    - The data was collected from 2017 to March 2020, directly matching the timeframe of the dataset instances. The data reflects the health statuses and conditions of the respondents during this specific period before the COVID-19 pandemic.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
    - Yes, the NHANES study protocols are reviewed by the NCHS Ethics Review Board (ERB) to ensure compliance with ethical standards and protection of participants. The ERB reviews all aspects of the NHANES, including data collection, consent processes, and the use and storage of data.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
    - Data was collected directly from the individuals through structured interviews and physical examinations conducted by NHANES personnel.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
    - Participants were fully informed about the nature of the study, the types of data collected, and the use of the data. They were provided with detailed consent forms that explained the data collection process and their rights as participants, including privacy protections and the voluntary nature of their participation.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
    - Yes, informed consent was obtained from all participants. The consent process involved participants reading the consent form and having the opportunity to ask questions. Consent was documented electronically in the CAPI system. The consent form detailed the participant's rights, the confidential handling of their data, and the purposes of the study.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
    - Participants were informed that they could withdraw from the study at any time without any penalty or loss of benefits to which they were otherwise entitled. The NHANES staff provided contact information for participants to use if they chose to withdraw their consent.

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
    - An impact analysis was conducted to assess the potential risks associated with data collection and use. The analysis considered privacy risks, data security, and the impact of data collection on participants. Measures were implemented to mitigate any identified risks, ensuring the protection of participant data and compliance with federal regulations.

12. *Any other comments?*
    - The NHANES team is committed to maintaining high ethical standards and transparency in its data collection and research processes. The study's methodology and results are regularly published in scientific journals and presented at conferences, contributing to public health knowledge and practice.

# Preprocessing/cleaning/labeling

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
   - Yes, preprocessing included the cleaning of data to correct common entry errors, such as the confusion of AM and PM in time entries and the transposition of digits. Instances with extreme or implausible sleep duration times were either corrected or marked as missing based on additional information from interviewer comments.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
   - The raw data, as collected, is maintained alongside the processed dataset to allow for validation of processing steps and to enable further analysis if required. Both versions of the dataset are available through the NHANES website.

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
   - The software tools used for data processing, primarily statistical and data management software supported by the CDC, are standardized and not specifically available for public use. However, the methodologies and procedures used are detailed in the NHANES documentation.

4. *Any other comments?*
   - The preprocessing steps are designed to ensure that the data is reliable and suitable for analysis, with a focus on maintaining the integrity and usability of the data for health research.

# Distribution

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
   - The dataset is publicly available and can be accessed by researchers, policymakers, and the public. It is intended to support a wide range of health-related research and policy-making.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
   - The dataset is distributed through the NHANES website, where users can download the data files directly. There is no DOI assigned to the dataset; however, each cycle of NHANES data is uniquely identified and can be cited using the specific cycle reference.

3. *When will the dataset be distributed?*
   - The dataset is already available for public access following its release in July 2021.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
   - The dataset is distributed under the terms of the Centers for Disease Control and Prevention's public use data policy, which allows for unrestricted use of the data for research and educational purposes. There are no fees associated with accessing the data.

5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
   - There are no third-party IP restrictions on the data. The NHANES data is fully controlled and distributed by the CDC, ensuring open access.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
   - No export controls or regulatory restrictions apply to the dataset. The data is intended for public use and is accessible worldwide.

7. *Any other comments?*
   - The NHANES program is committed to providing high-quality, accessible data that can inform health research and policy. The program encourages the use of its data to advance public health knowledge and practices globally.

# Maintenance

1. *Who will be supporting/hosting/maintaining the dataset?*
   - The dataset is supported, hosted, and maintained by the National Center for Health Statistics (NCHS), which is part of the Centers for Disease Control and Prevention (CDC).

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
   - The NCHS can be contacted through their website, where users can find contact information for various departments and services.

3. *Is there an erratum? If so, please provide a link or other access point.*
   - Updates, corrections, and errata related to the NHANES data are posted on the NHANES website. Users can access this information to ensure they are using the most accurate and up-to-date data.

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
   - The NHANES data is updated periodically as new data collection cycles are completed. Updates are managed by the NCHS and are communicated through the NHANES website and through official CDC communications.

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
   - The data collected in NHANES is retained indefinitely as part of the historical and scientific record of population health. Data retention policies are governed by federal regulations and guidelines, which ensure the protection of participant information and the long-term availability of the data for research.

6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
   - Older versions of the NHANES data are maintained and remain accessible to the public. This ensures that researchers can access historical data for longitudinal studies and comparisons across different time periods.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
   - While external contributions to the dataset itself are not typically solicited, the NHANES program encourages researchers to use the data in their studies and to publish their findings. Contributions to the broader scientific community through research and publications based on NHANES data are welcomed and supported by the NCHS.

8. *Any other comments?*
   - The NHANES dataset is a critical resource for health research in the United States and globally. It provides invaluable insights into the health status, lifestyle, and risk factors of the U.S. population and serves as a foundation for public health policy and interventions.

\newpage

# References

### 2024-04-18 scripts/00_environment-setup.R
added shiny app

@@ -9,7 +9,7 @@
 required_packages <- c(
   "ggplot2", "dplyr", "tidyverse", "here", "readr", 
   "knitr", "kableExtra", "rstanarm", "haven", "testthat",
-  "arrow", "gridExtra", "modelsummary", "bayesplot"
+  "arrow", "gridExtra", "modelsummary", "bayesplot", "shiny"
 )
 
 # Function to check and install any missing packages

### 2024-04-18 scripts/06_shiny-app.R

@@ -1,0 +1,84 @@
+#### Preamble ####
+# Purpose: To provide an interactive tool for predicting the likelihood of sleep troubles based on individual sleep patterns.
+# Author: Terry Tu
+# Date: 17 April 2024
+# Contact: xiangyu.tu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'shiny', 'dplyr', 'shinyjs'
+
+# Load required libraries
+library(shiny)
+library(dplyr)
+library(shinyjs)  # for using JavaScript to dynamically change styles
+
+# Define UI
+ui <- fluidPage(
+  useShinyjs(),  # Initialize shinyjs
+  titlePanel("Sleep Trouble Prediction Tool"),
+  tags$p("This interactive tool allows users to input their own sleep patterns to predict the likelihood of reporting sleep troubles. Based on a logistic regression model, the application estimates the probability that an individual might report sleep issues to a healthcare provider. To learn more about the data and model, visit the GitHub repository: https://github.com/TEJMaster/Sleep-Disorder-Analysis."),
+  sidebarLayout(
+    sidebarPanel(
+      numericInput("weekdaySleep", "Weekday Sleep Duration (hours)", value = 8, min = 2, max = 14),
+      numericInput("weekendSleep", "Weekend Sleep Duration (hours)", value = 8, min = 2, max = 14),
+      selectInput("snoringFrequency", "How often do you snore?",
+                  choices = c("Never" = 0, 
+                              "Rarely - 1-2 nights a week" = 1, 
+                              "Occasionally - 3-4 nights a week" = 2, 
+                              "Frequently - 5 or more nights a week" = 3)),
+      selectInput("breathingPause", "How often do you snort or stop breathing",
+                  choices = c("Never" = 0, 
+                              "Rarely - 1-2 nights a week" = 1, 
+                              "Occasionally - 3-4 nights a week" = 2, 
+                              "Frequently - 5 or more nights a week" = 3)),
+      selectInput("overlySleepy", "How often feel overly sleepy during day?",
+                  choices = c("Never" = 0, 
+                              "Rarely - 1 time a month" = 1, 
+                              "Sometimes - 2-4 times a month" = 2, 
+                              "Often - 5-15 times a month" = 3, 
+                              "Almost always - 16-30 times a month" = 4)),
[…]

### 2024-04-19 paper/paper.qmd
added shiny app

@@
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-cap: "Examining how the model fits, and is affected by the data"
 #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
@@
 
+## C Shiny App {.unnumbered}
+
+```{r fig-shiny-app, fig.cap="Demo of the shiny application", out.width='70%', out.height='70%', echo=FALSE}
+knitr::include_graphics(here("other", "sketches", "shiny-app.jpg"))
+```
+
+The Shiny app presented in @fig-shiny-app is a user-friendly tool designed to help individuals understand their risk of having sleep troubles based on their self-reported sleep habits.
+Utilizing a pre-trained logistic regression model from this project, this application allows users to input various factors related to their sleep, such as weekday and weekend sleep duration, frequency of snoring, incidents of breathing pauses, and levels of daytime sleepiness.
+
+To use the app, simply navigate through the input fields, providing accurate information about your regular sleep patterns.
+You can specify how many hours you typically sleep on weekdays and weekends, as well as how often you experience snoring, breathing pauses while sleeping, and excessive sleepiness during the day.
+These categories have predefined options ranging from 'Never' to 'Frequently' for snoring and breathing interruptions, and from 'Never' to 'Almost always' for feeling overly sleepy, making it straightforward to match your experiences to the app’s metrics.
+
+After inputting your data, click on the "Predict Probability" button.
+The app will process your inputs through the logistic regression model and return a probability score.
+This score represents your estimated likelihood of experiencing sleep troubles that you might report to a healthcare provider.
+The outcome is presented in a clear and concise manner at the bottom of the app, with the probability score highlighted in a color corresponding to the severity of the prediction: green for low, yellow for moderate, and red for high probability.
+
+You can access the Sleep Trouble Prediction Tool (Shiny app) using [https://3tbrto-terry0tu.shinyapps.io/shiny/](https://3tbrto-terry0tu.shinyapps.io/shiny/) to explore your sleep patterns and their implications for your wellbeing.
+
+\newpage
+
 # References

### 2024-04-19 scripts/00_environment-setup.R

@@ -7,7 +7,7 @@
 
 # Define a vector of required packages
 required_packages <- c(
-  "ggplot2", "dplyr", "tidyverse", "here", "readr", 
+  "ggplot2", "dplyr", "tidyverse", "here", "readr", "shinyjs",
   "knitr", "kableExtra", "rstanarm", "haven", "testthat",
   "arrow", "gridExtra", "modelsummary", "bayesplot", "shiny"
 )

### 2024-04-19 paper/paper.qmd
grammar fixes

@@
 
-Utilizing the National Health and Nutrition Examination Survey (NHANES) dataset from 2017 to March 2020, our research examines how various sleep behaviors—such as snoring frequency, daytime sleepiness, and sleep regularity—correlate with the likelihood of reporting sleep issues to healthcare providers.
+Utilizing the National Health and Nutrition Examination Survey (NHANES) dataset from 2017 to March 2020 [@Prevention_2021], our research examines how various sleep behaviors—such as snoring frequency, daytime sleepiness, and sleep regularity—correlate with the likelihood of reporting sleep issues to healthcare providers.
 The primary estimand of our analysis is the association between these behaviors and the probability of an individual reporting sleep disturbances, quantified through logistic regression.
@@
 
-The dataset underpinning this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020.
+The dataset used in this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020.
 This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially.
-After meticulous data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional.
+After some data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional.
 The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].
@@
 The survey participants' ages range widely, reflecting the diversity of the American population.
-Variables are finely tuned to capture the multifaceted nature of sleep, encompassing aspects such as duration, disruptions, and subjective experiences of daytime sleepiness.
-The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
+Variables are finely tuned to capture the multifaceted nature of sleep, including aspects such as duration, disruptions, and subjective experiences of daytime sleepiness.
+The NHANES protocol ensures that this dataset is a robust and reliable source of information, assure ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
 For further information on the data cleaning specifics and validation checks, please see the supplementary material in Appendix.
@@
 
-Our statistical exploration was conducted within the R programming environment [@r], leveraging its comprehensive ecosystem for data analysis.
-We utilized the tidyverse collection of R packages [@rTidyverse] to streamline our data processing tasks.
-The ggplot2 package [@rGgplot2] was instrumental in crafting insightful visualizations that articulated the intricate relationships within our data.
-The dplyr package [@rDplyr] provided a syntax that facilitated the manipulation and transformation of our dataset, enabling us to prepare the data effectively for logistic regression analysis.
-Data importation was efficiently handled by the readr package [@rReadr], known for its quick and user-friendly approach to reading tabular data.
-Navigational simplicity within our project's directories was achieved with the here package [@R-here], which reliably managed file paths without the need for manual path setting.
-The reproducibility of our research was enhanced by the knitr package [@rKnitr], which seamlessly wove R code into our report, ensuring that our findings are transparent and replicable.
-For tabular data presentation, kableExtra[@R-kableExtra] offered a suite of customization options that enhanced the readability and aesthetic appeal of our tables.
-The logistic regression model was developed using core functions in R, which provide robust methods for estimating the effects of various predictors on a binary outcome.
+Our analytical approach was completed using R [@r], utilizing the tidyverse [@rTidyverse] for data manipulation and ggplot2 [@rGgplot2] for visualization.
+The readr [@rReadr], here [@R-here], and arrow [@rArrow] packages were pivotal in handling data importation and directory management.
+Reproducible reporting was facilitated by knitr [@rKnitr] and kableExtra [@R-kableExtra] for enhancing table presentations.
+Bayesian logistic regression modeling was conducted with rstanarm [@rRstanarm], while model diagnostics were visualized using bayesplot [@rBayesplot].
+These tools, which formed the backbone for this paper as well as the scripts, ensured efficient and reproducible analyses, culminating in a robust set of findings.
 
@@
 
+\newpage
 ## Measurement:
@@
 This section is designed to provide a comprehensive understanding of the origins, processing, and analytical framework applied to each variable relevant to this study.
+
+\newpage
 
@@
 Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise.
-The predictors include weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$).
+The predictors include continous variables like weekday sleep duration ($x_{i1}$), and weekend sleep duration ($x_{i2}$); ordinal categorical variables such as snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and overly sleep frequency ($x_{i5}$).
 The logistic regression model is formulated as follows:
@@
 The priors for the intercept ($\beta_0$) and the slopes are deliberately chosen to reflect initial neutrality or skepticism about each predictor's impact on the likelihood of reporting sleep troubles.
-Specifically, $\beta_0$, $\beta_1$, and $\beta_2$ are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data, acknowledging that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.
+Specifically, $\beta_0$ for intercept, $\beta_1$, and $\beta_2$ for the continous variables about sleep hours are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data, acknowledging that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.
 
-For $\beta_3$, $\beta_4$, and $\beta_5$, which relate to snoring frequency, breathing pause frequency, and overly sleep frequency, the same Normal(0, 1) priors are used.
+For $\beta_3$, $\beta_4$, and $\beta_5$, which relate to ordinal categorical variables such as snoring frequency, breathing pause frequency, and overly sleep frequency are set to have the same Normal(0, 1) priors.
 This choice indicates no strong preconceived notions about the direction or magnitude of their impact, allowing the data itself to guide the inference about these predictors without the influence of strong initial assumptions.
@@
 Our Bayesian logistic regression model is designed to investigate the associations between various sleep-related factors and the likelihood of reporting sleep troubles.
-The priors for the coefficients are set with specific hypotheses in mind, informed by existing literature and plausible biological mechanisms:
+The priors for the coefficients are set with specific hypotheses in mind, informed by knowleged gained from the data exploration section:
 
@@
 
-## Model Equation
+## Logistic Regression Model for Predicting Individual Sleep Disorder Probability
 
@@
 
-## Predict Sleep Disorder Through Self Reported Patterns
+## Example of Predict Sleep Disorder Through Self Reported Patterns
 

### 2024-04-19 scripts/03_test-data-validity.R

@@ -4,17 +4,18 @@
 # Date: 2 April 2024
 # Contact: xiangyu.tu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 'dplyr', 'here', 'readr', 'testthat'
+# Pre-requisites: 'dplyr', 'here', 'haven', 'arrow', 'testthat'
 # The cleaned dataset must be located in the 'data/analysis_data' directory before running this script
 
 #### Workspace setup ####
-library(readr)
+library(haven)
+library(arrow)
 library(testthat)
 library(dplyr)
 library(here)
 
 # Load the cleaned data
-cleaned_data <- read_csv(here("data", "analysis_data", "cleaned_data.csv"))
+cleaned_data <- read_parquet(here("data", "analysis_data", "cleaned_data.parquet"))
 
 # Test 1: Check if every column has a valid integer ID (SEQN)
 test_that("Every column has a valid integer ID", {
